### 2022-04-12 analysis/tgispine_hypotheses.Rmd
using pilot to inform simulation, creating pilot plots

---
title: "TGI-spine"
author: "Francesca Fardo"
date: "18/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

## Simulate results for one participant
```{r}
simulate_subj <- function(subid) {
trialN = 9
conditionN = 8
ratingN = 3
df <- data.frame(matrix(ncol = 0, nrow = trialN*conditionN*ratingN))
df$id <- rep(subid,trialN*conditionN*ratingN)
df$trialn <- rep(1:trialN,conditionN*ratingN)
df$dermatome <- rep(rep(c(rep("within",trialN*2), rep("across",trialN*2)),2),ratingN)
df$cold_location <- rep(rep(c(rep("dist_rostr",trialN), rep("prox_caud", trialN)),conditionN/2),ratingN)
df$stimuli <- rep(c(rep("tgi",trialN*conditionN/2), rep("non-tgi",trialN*conditionN/2)),ratingN)
df$quality <- c(rep("cold",trialN*conditionN), rep("warm",trialN*conditionN), rep("burn",trialN*conditionN))

# The following intervals and mean can be changed to reflect hypotheses
# Cold ratigs: simulation
w_dr_tgi_cold <- rnorm(trialN, mean=sample(5:25,1), sd=5)
w_pc_tgi_cold <- rnorm(trialN, mean=sample(5:25,1), sd=5)
a_dr_tgi_cold <- rnorm(trialN, mean=sample(5:50,1), sd=5)
a_pc_tgi_cold <- rnorm(trialN, mean=sample(5:40,1), sd=5)
w_dr_nontgi_cold <- rnorm(trialN, mean=sample(10:60,1), sd=15)
w_pc_nontgi_cold <- rnorm(trialN, mean=sample(10:60,1), sd=15)
a_dr_nontgi_cold <- rnorm(trialN, mean=sample(10:55,1), sd=15)
a_pc_nontgi_cold <- rnorm(trialN, mean=sample(10:55,1), sd=15)

# Warm ratigs: simulation
w_dr_tgi_warm <- rnorm(trialN, mean=sample(10:65,1), sd=15)
w_pc_tgi_warm <- rnorm(trialN, mean=sample(10:70,1), sd=15)
a_dr_tgi_warm <- rnorm(trialN, mean=sample(5:60,1), sd=15)
a_pc_tgi_warm <- rnorm(trialN, mean=sample(5:60,1), sd=15)
w_dr_nontgi_warm <- rnorm(trialN, mean=sample(5:25,1), sd=10)
w_pc_nontgi_warm <- rnorm(trialN, mean=sample(5:25,1), sd=10)
a_dr_nontgi_warm <- rnorm(trialN, mean=sample(5:25,1), sd=10)
a_pc_nontgi_warm <- rnorm(trialN, mean=sample(5:25,1), sd=10)

# Burning ratings: simulation
 w_dr_tgi_burn <- rnorm(trialN, mean=sample(10:65,1), sd=15)
 w_pc_tgi_burn <- rnorm(trialN, mean=sample(10:60,1), sd=15)
 a_dr_tgi_burn <- rnorm(trialN, mean=sample(10:70,1), sd=15)
 a_pc_tgi_burn <- rnorm(trialN, mean=sample(5:60,1), sd=15)
 w_dr_nontgi_burn <- rnorm(trialN, mean=sample(0:10,1), sd=5)
 w_pc_nontgi_burn <- rnorm(trialN, mean=sample(0:10,1), sd=5)
 a_dr_nontgi_burn <- rnorm(trialN, mean=sample(0:10,1), sd=5)
 a_pc_nontgi_burn <- rnorm(trialN, mean=sample(0:10,1), sd=5)

df$rating <- c(w_dr_tgi_cold, w_pc_tgi_cold, a_dr_tgi_cold, a_pc_tgi_cold, w_dr_nontgi_cold, w_pc_nontgi_cold, a_dr_nontgi_cold, a_pc_nontgi_cold,
               w_dr_tgi_warm, w_pc_tgi_warm, a_dr_tgi_warm, a_pc_tgi_warm, w_dr_nontgi_warm, w_pc_nontgi_warm, a_dr_nontgi_warm, a_pc_nontgi_warm,
               w_dr_tgi_burn, w_pc_tgi_burn, a_dr_tgi_burn, a_pc_tgi_burn, w_dr_nontgi_burn, w_pc_nontgi_burn, a_dr_nontgi_burn, a_pc_nontgi_burn)
return(df)}
```

# Simulate results for 30 participants
```{r}
df30 <- data.frame()
for (i in 1:60) {
  df <- simulate_subj(i)
  df30 <- rbind(df30,df)}
df30
```
```{r}
df30$conditionN[df$dermatome == "within" & df30$cold_location == "dist_rostr"] <- 1
df30$conditionN[df$dermatome == "within" & df30$cold_location == "prox_caud"] <- 2
df30$conditionN[df$dermatome == "across" & df30$cold_location == "dist_rostr"] <- 3
df30$conditionN[df$dermatome == "across" & df30$cold_location == "prox_caud"] <- 4
df30
```

## Summary stats: subject-level
```{r}
# Subject-level
df <- df30
df_subj <- aggregate(df$rating, by = list(id = df$id, conditionN = df$conditionN, dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = mean)
#df_subj$dermatome <- factor(df_subj$dermatome, levels = c('within','across')) # re-order conditions
df_subj
```
## Summary stats: group-level
```{r}
# Group-level stats
df <- df_subj
group_mean   <- aggregate(df$x, by = list(conditionN = df$conditionN, dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = mean)
group_median <- aggregate(df$x, by = list(conditionN = df$conditionN,dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = median)
group_sd     <- aggregate(df$x, by = list(conditionN = df$conditionN,dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = sd)
group_se     <- group_sd$x/sqrt(length(unique(df$id)))

df_group <- data.frame(group_mean, mean = group_mean$x, median = group_median$x, sd = group_sd$x, se = group_se)
df_group
```

## Settings for plots

```{r, echo = FALSE}
# Load libraries
library("ggplot2")
library("gghalves")
library("RColorBrewer")

# Define colors
reds <- brewer.pal(6, "Reds")
blues <- brewer.pal(6, "Blues")
purps <- brewer.pal(6, "Purples")

col_cold_n <- blues[5]
col_cold_w <- blues[3]
col_warm_n <- reds[3]
col_warm_w <- reds[5]
col_burn_n <- purps[3]
col_burn_w <- purps[5]

# Parameters for plots
tgi_theme = theme( 
    text = element_text(size = 14),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title=element_text(size=12), 
    legend.text=element_text(size=12),
    plot.title = element_text(lineheight=.8, face="bold", size = 12),
    panel.border = element_rect(size=1),
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
    legend.key = element_blank(),
    axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))

# Other parameters
nudge1 = .2
nudge2 = .3
nudge3 = .4

# Define how to plot half violins
tgi_half_violin <- function(df_subj, tgi_dermatome, tgi_cold_location, tgi_stimuli, nudge, color) {
v1 <- geom_half_violin(
   data = df_subj %>% filter(dermatome == tgi_dermatome & cold_location == tgi_cold_location & stimuli == tgi_stimuli), 
   aes(x = conditionN, y = x), position = position_nudge(x = nudge), side = "r", fill = color, alpha = .8)
return(v1)}

# Define how to plot group means and error bars
tgi_group_mean <- function(df_group, tgi_dermatome, tgi_cold_location, tgi_stimuli, nudge, color, shape) {
m1 <- geom_point(
  data = df_group %>% filter(dermatome == tgi_dermatome & cold_location == tgi_cold_location & stimuli == tgi_stimuli), 
  aes(x = conditionN, y = mean), position = position_nudge(x = nudge + .05), size = 3, shape = shape)
return(m1)}

# Define how to plot error bars
tgi_group_error <- function(df_group, tgi_dermatome, tgi_cold_location, tgi_stimuli, tgi_nudge, tgi_color) {
e1 <- geom_errorbar(
  data = df_group %>% filter(dermatome == tgi_dermatome & cold_location == tgi_cold_location & stimuli == tgi_stimuli), 
  aes(x = conditionN, y = mean, ymin = mean - se, ymax = mean + se), position = position_nudge(tgi_nudge + .05), width = 0)
return(e1)}

```

## Plot cold ratings

```{r, echo=FALSE}
# Subset of data corresponding to cold ratings
df_subj_cold  <- subset(df_subj, quality == "cold")
df_group_cold <- subset(df_group, quality == "cold")

# Set jitter for geom_point
set.seed(321)
df_subj_cold$xj <- jitter(df_subj_cold$conditionN, amount=.1)

c1 <- ggplot(data = df_subj_cold, mapping = aes(x = conditionN, y = x, fill = stimuli, shape = stimuli, color = stimuli)) +

   # Add theme
   theme_bw()+
   tgi_theme +
   geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5) +

   # Add points (for each subject and condition)
   geom_point(aes(x = xj), size = 2, alpha = .8) +

   # Add half violin plots
   tgi_half_violin(df_subj_cold, "within", "dist_rostr", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "within", "prox_caud", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "across", "dist_rostr", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "across", "prox_caud", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "within", "dist_rostr", "tgi", nudge1, col_cold_w) +
   tgi_half_violin(df_subj_cold, "within", "prox_caud", "tgi", nudge1, col_cold_w) +
   tgi_half_violin(df_subj_cold, "across", "dist_rostr", "tgi", nudge1, col_cold_w) +
   tgi_half_violin(df_subj_cold, "across", "prox_caud", "tgi", nudge1, col_cold_w) +

   # Add group summary stats (between-condition lines, errors bars, and means)
   geom_line(data = df_group_cold %>% filter(stimuli == "non-tgi"), aes(x = conditionN + .45, y = mean)) +
   geom_line(data = df_group_cold %>% filter(stimuli == "tgi"), aes(x = conditionN + .35, y = mean)) +

   tgi_group_error(df_group_cold, "within", "dist_rostr", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "within", "prox_caud", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "across", "dist_rostr", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "across", "prox_caud", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "within", "dist_rostr", "tgi", nudge2, col_cold_w) +
   tgi_group_error(df_group_cold, "within", "prox_caud", "tgi", nudge2, col_cold_w) +
   tgi_group_error(df_group_cold, "across", "dist_rostr", "tgi", nudge2, col_cold_w) +
   tgi_group_error(df_group_cold, "across", "prox_caud", "tgi", nudge2, col_cold_w) +

   tgi_group_mean(df_group_cold, "within", "dist_rostr", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "within", "prox_caud", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "across", "dist_rostr", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "across", "prox_caud", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "within", "dist_rostr", "tgi", nudge2, col_cold_w, 21) +
   tgi_group_mean(df_group_cold, "within", "prox_caud", "tgi", nudge2, col_cold_w, 21) +
   tgi_group_mean(df_group_cold, "across", "dist_rostr", "tgi", nudge2, col_cold_w, 21) +
   tgi_group_mean(df_group_cold, "across", "prox_caud", "tgi", nudge2, col_cold_w, 21) +

   # Define additional settings
   scale_shape_manual(values = c(22, 21)) +
   scale_x_continuous(breaks=c(1.2, 2.2, 3.2, 4.2), labels=c("distal", "proximal", "rostral", "caudal"), limits=c(0.6, 5)) +
   scale_fill_manual(labels = c("non-TGI", "TGI"), values=c(blues[6],blues[3])) +
   scale_color_manual(labels = c("non-TGI", "TGI"), values=c(blues[6], "gray25")) +
   guides(shape = FALSE) +
   labs(title = "", x = "", y = "Cold ratings", fill = "Stimuli", color = "Stimuli") +
   ggtitle('Figure 1: Cold perception')

c1
```
## Plot warm ratings

```{r, echo=FALSE}
# Subset of data corresponding to warm ratings
df_subj_warm  <- subset(df_subj, quality == "warm")
df_group_warm <- subset(df_group, quality == "warm")

# Set jitter for geom_point
set.seed(321)
df_subj_warm$xj <- jitter(df_subj_warm$conditionN, amount=.1)

w1 <- ggplot(data = df_subj_warm, mapping = aes(x = conditionN, y = x, fill = stimuli, shape = stimuli, color = stimuli)) +

   # Add theme
   theme_bw()+
   tgi_theme +
   geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5) +

   # Add points (for each subject and condition)
   geom_point(aes(x = xj), size = 2, alpha = .8) +

   # Add half violin plots
   tgi_half_violin(df_subj_warm, "within", "dist_rostr", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "within", "prox_caud", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "across", "dist_rostr", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "across", "prox_caud", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "within", "dist_rostr", "tgi", nudge1, col_warm_w) +
   tgi_half_violin(df_subj_warm, "within", "prox_caud", "tgi", nudge1, col_warm_w) +
   tgi_half_violin(df_subj_warm, "across", "dist_rostr", "tgi", nudge1, col_warm_w) +
   tgi_half_violin(df_subj_warm, "across", "prox_caud", "tgi", nudge1, col_warm_w) +

   # Add group summary stats (between-condition lines, errors bars, and means)
   geom_line(data = df_group_warm %>% filter(stimuli == "non-tgi"), aes(x = conditionN + .45, y = mean)) +
   geom_line(data = df_group_warm %>% filter(stimuli == "tgi"), aes(x = conditionN + .35, y = mean)) +

   tgi_group_error(df_group_warm, "within", "dist_rostr", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "within", "prox_caud", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "across", "dist_rostr", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "across", "prox_caud", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "within", "dist_rostr", "tgi", nudge2, col_warm_w) +
   tgi_group_error(df_group_warm, "within", "prox_caud", "tgi", nudge2, col_warm_w) +
   tgi_group_error(df_group_warm, "across", "dist_rostr", "tgi", nudge2, col_warm_w) +
   tgi_group_error(df_group_warm, "across", "prox_caud", "tgi", nudge2, col_warm_w) +

   tgi_group_mean(df_group_warm, "within", "dist_rostr", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "within", "prox_caud", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "across", "dist_rostr", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "across", "prox_caud", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "within", "dist_rostr", "tgi", nudge2, col_warm_w, 21) +
   tgi_group_mean(df_group_warm, "within", "prox_caud", "tgi", nudge2, col_warm_w, 21) +
   tgi_group_mean(df_group_warm, "across", "dist_rostr", "tgi", nudge2, col_warm_w, 21) +
   tgi_group_mean(df_group_warm, "across", "prox_caud", "tgi", nudge2, col_warm_w, 21) +

   # Define additional settings
   scale_shape_manual(values = c(22, 21)) +
   scale_x_continuous(breaks=c(1.2, 2.2, 3.2, 4.2), labels=c("distal", "proximal", "rostral", "caudal"), limits=c(0.6, 5)) +
   scale_fill_manual(labels = c("non-TGI", "TGI"), values=c(col_warm_n, col_warm_w)) +
   scale_color_manual(labels = c("non-TGI", "TGI"), values=c("gray25", reds[6])) +
   guides(shape = FALSE) +
   labs(title = "", x = "", y = "Warm ratings", fill = "Stimuli", color = "Stimuli") +
   ggtitle('Figure 2: Warm perception')

w1
```
## Plot burn ratings

```{r, echo=FALSE}
# Subset of data corresponding to warm ratings
df_subj_burn  <- subset(df_subj, quality == "burn")
df_group_burn <- subset(df_group, quality == "burn")

# Set jitter for geom_point
set.seed(321)
df_subj_burn$xj <- jitter(df_subj_burn$conditionN, amount=.1)

b1 <- ggplot(data = df_subj_burn, mapping = aes(x = conditionN, y = x, fill = stimuli, shape = stimuli, color = stimuli)) +

   # Add theme
   theme_bw()+
   tgi_theme +
   geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5) +

   # Add points (for each subject and condition)
   geom_point(aes(x = xj), size = 2, alpha = .8) +

   # Add half violin plots
   tgi_half_violin(df_subj_burn, "within", "dist_rostr", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "within", "prox_caud", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "across", "dist_rostr", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "across", "prox_caud", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "within", "dist_rostr", "tgi", nudge1, col_burn_w) +
   tgi_half_violin(df_subj_burn, "within", "prox_caud", "tgi", nudge1, col_burn_w) +
   tgi_half_violin(df_subj_burn, "across", "dist_rostr", "tgi", nudge1, col_burn_w) +
   tgi_half_violin(df_subj_burn, "across", "prox_caud", "tgi", nudge1, col_burn_w) +

   # Add group summary stats (between-condition lines, errors bars, and means)
   geom_line(data = df_group_burn %>% filter(stimuli == "non-tgi"), aes(x = conditionN + .45, y = mean)) +
   geom_line(data = df_group_burn %>% filter(stimuli == "tgi"), aes(x = conditionN + .35, y = mean)) +

   tgi_group_error(df_group_burn, "within", "dist_rostr", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "within", "prox_caud", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "across", "dist_rostr", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "across", "prox_caud", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "within", "dist_rostr", "tgi", nudge2, col_burn_w) +
   tgi_group_error(df_group_burn, "within", "prox_caud", "tgi", nudge2, col_burn_w) +
   tgi_group_error(df_group_burn, "across", "dist_rostr", "tgi", nudge2, col_burn_w) +
   tgi_group_error(df_group_burn, "across", "prox_caud", "tgi", nudge2, col_burn_w) +

   tgi_group_mean(df_group_burn, "within", "dist_rostr", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "within", "prox_caud", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "across", "dist_rostr", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "across", "prox_caud", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "within", "dist_rostr", "tgi", nudge2, col_burn_w, 21) +
   tgi_group_mean(df_group_burn, "within", "prox_caud", "tgi", nudge2, col_burn_w, 21) +
   tgi_group_mean(df_group_burn, "across", "dist_rostr", "tgi", nudge2, col_burn_w, 21) +
   tgi_group_mean(df_group_burn, "across", "prox_caud", "tgi", nudge2, col_burn_w, 21) +

   # Define additional settings
   scale_shape_manual(values = c(22, 21)) +
   scale_x_continuous(breaks=c(1.2, 2.2, 3.2, 4.2), labels=c("distal", "proximal", "rostral", "caudal"), limits=c(0.6, 5)) +
   scale_fill_manual(labels = c("non-TGI", "TGI"), values=c(col_burn_n, col_burn_w)) +
   scale_color_manual(labels = c("non-TGI", "TGI"), values=c("gray25", purps[6])) +
   guides(shape = FALSE) +
   labs(title = "", x = "", y = "Burn ratings", fill = "Stimuli", color = "Stimuli") +
   ggtitle('Figure 3: Burn perception')

b1
```

# running inferential stats on simulated data - to gain idea of effect size

```{r}
library(ez)
library(psychReport)
# isolate just TGI for purposes of determining effect size
df_tgi <- df_subj[df_subj$stimuli == 'tgi' ,]

# burning VAS
burn_anova <- ezANOVA(
  data = df_tgi[df_tgi$quality == 'burn' ,]
  , dv = .(x)
  , wid = .(id)
  , within = .(dermatome, cold_location),
  return_aov = TRUE,
  detailed = TRUE
)

# warm VAS
warm_anova <- ezANOVA(
  data = df_tgi[df_tgi$quality == 'warm' ,]
  , dv = .(x)
  , wid = .(id)
  , within = .(dermatome, cold_location),
  return_aov = TRUE,
  detailed = TRUE
)

# cold VAS
cold_anova <- ezANOVA(
  data = df_tgi[df_tgi$quality == 'cold' ,]
  , dv = .(x)
  , wid = .(id)
  , within = .(dermatome, cold_location),
  return_aov = TRUE,
  detailed = TRUE
)

## t-test
# create seperate data-frames to allow for comparison
within <- df_subj[df_subj$dermatome=='within' ,]
across <- df_subj[df_subj$dermatome=='across' ,]
proxc <- within[within$cold_location== 'prox_caud' ,]
distc <- within[within$cold_location=='dist_rostr' ,]
caudc <- across[across$cold_location=='prox_caud' ,]
rostc <- across[across$cold_location=='dist_rostr' ,]

## H1
names(within)[7] <- 'within'
names(across)[7] <- 'across'
# merge
H1_df <- merge(within, across, by = c('id','cold_location','stimuli','quality'))
# difference
H1_df$diff <- H1_df$within - H1_df$across

H1test <- t.test(H1_df$within, H1_df$across, alternative = 'two.sided', paired = TRUE)

```