### 2022-04-22 analysis/tgispine_og-plots-exp1.Rmd
adding tgispine analysis code for old data - to use for power calculation

---
title: "TGI plots"
output: html_notebook
---

# Libraries
```{r warning = FALSE}
library(readr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(dplyr)
```

# Import data 
```{r}
path <- '/Users/au706616/Documents/Experiments/SPINALTGI/PreviousData/'
filename <- 'tgi-spine2-data-exp1.csv'
df <- read_delim(file.path(path,filename), ",", col_names = TRUE)
df$dermatome <- factor(df$dermatome, levels = c('within','across'))
df$location_arm <- factor(df$location_arm, levels = c('forearm','upperarm'))
df$location_cold <- factor(df$location_cold, levels = c('dist_caud','prox_rost'))
df$id <- factor(df$id)
df
```

# Summary
```{r}
n_stim <- df %>% count(id = df$id, dermatome = df$dermatome, location_cold = df$location_cold, stim = df$stim) # number of stimuli per condition
df_median <- aggregate(list(cold = df$cold, warm = df$warm, pain = df$pain), by = list(id = df$id, dermatome = df$dermatome, location_cold = df$location_cold), FUN = median)
df_median
```
# Analysis and effect sizes
```{r}
# cold ratings anova

# cold ratings linear mixed model
# warm ratings anova
# warm ratings linear mixed model
```

# Theme for plots 
```{r warning = FALSE}
variance_theme = theme( 
    text = element_text(size = 14),
    #axis.line = element_line(size = 0.7, color = "black"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title=element_text(size=12), 
    legend.text=element_text(size=12),
    plot.title = element_text(lineheight=.8, face="bold", size = 12),
    panel.border = element_rect(size=1),
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
    legend.key = element_blank(),
    axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))
    # axis.text.x = element_text()

# set a figure directory
#figdir<-"~/Dropbox/!Projects/2017_TGI_ginty/figure/"
# dpi resolution
resolution = 600
```
# Figure 1: cold ratings
```{r}
p1 <-ggplot(df_median, aes(x = interaction(location_cold,dermatome), y = cold, fill = dermatome))+
   theme_bw()+
   variance_theme+
   #geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5)+
   geom_boxplot(outlier.shape = NA, notch = FALSE)+ 
   geom_point(aes(x = interaction(location_cold,dermatome), fill = dermatome), shape = 21, size = 3)+
   scale_x_discrete(labels=c("Distal", "Proximal", "Caudal", "Rostral"))+
   scale_y_continuous(limits = c(0,100))+
   guides(fill = FALSE)+
   scale_fill_brewer(palette = "Blues", direction = 1) +
   theme(legend.position="none")+
   ylab("Cold Ratings")+
   xlab("")
p1

# ggsave(file=paste(figdir, "EXP1_tgi_boxplot_cold.pdf"), width=7, height=7, dpi = resolution, units = "cm")

```

# Figure 2: warm ratings
```{r}
p2 <-ggplot(df_median, aes(x = interaction(location_cold,dermatome), y = warm, fill = dermatome))+
   theme_bw()+
   variance_theme+
   #geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5)+
   geom_boxplot(outlier.shape = NA, notch = FALSE)+ 
   geom_point(aes(x = interaction(location_cold,dermatome), fill = dermatome), shape = 21, size = 3)+
   scale_x_discrete(labels=c("Distal", "Proximal", "Caudal", "Rostral"))+
   scale_y_continuous(limits = c(0,100))+
   guides(fill = FALSE)+
   scale_fill_brewer(palette = "Reds", direction = -1) +
   theme(legend.position="none")+
   ylab("Warm Ratings")+
   xlab("")
p2

#ggsave(file=paste(figdir, "EXP1_tgi_boxplot_warm.pdf"), width=7, height=7, dpi = resolution, units = "cm")

```
# Figure 3: pain ratings
```{r}
p3 <-ggplot(df_median, aes(x = interaction(location_cold,dermatome), y = pain, fill = dermatome))+
   theme_bw()+
   variance_theme+
   #geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5)+
   geom_boxplot(outlier.shape = NA, notch = FALSE)+ 
   geom_point(aes(x = interaction(location_cold,dermatome), fill = dermatome), shape = 21, size = 3)+
   scale_x_discrete(labels=c("Distal", "Proximal", "Caudal", "Rostral"))+
   scale_y_continuous(limits = c(0,100))+
   guides(fill = FALSE)+
   scale_fill_brewer(palette = "Purples", direction = -1) +
   theme(legend.position="none")+
   ylab("Pain Ratings")+
   xlab("")
p3

#ggsave(file=paste(figdir, "EXP1_tgi_boxplot_pain.pdf"), width=7, height=7, dpi = resolution, units = "cm")
```