### 2022-05-10 analysis/tgispine_lme_power.Rmd
creating density plots

---
title: "TGI-spinal-power"
author: "A.G. Mitchell"
date: '2022-04-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(simr)
library(broom)
library(brms)
library(modelsummary)
library(tibble)
library(DHARMa)
library(glmmTMB)
library(reshape2)
library(tidyverse)
library(plyr)

set.seed(1234)  # Make everything reproducible

# Define the goodness-of-fit stats to include in modelsummary()
gof_stuff <- tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "N", 0,
  "r.squared", "R²", 3
)

# setting working directory
datPath <- '/Users/au706616/Documents/Experiments/SPINALTGI/Raw/'
knitr::opts_knit$set(root.dir = datPath)
```

# Getting data and reformatting to fit model
```{r}
# import compiled data
df <- read.csv('STGI_compiled-data.csv')

# need to recode levels for cold_probe to reduce number of levels from 4 to 2
df$cold_location[df$cold_probe == 'distal'] <- 'dist_rostr' 
df$cold_location[df$cold_probe == 'rostral'] <- 'dist_rostr'
df$cold_location[df$cold_probe == 'proximal'] <- 'prox_caud'
df$cold_location[df$cold_probe == 'caudal'] <- 'prox_caud'
df$cold_location <- as.factor(df$cold_location)

# reshape data so that there is one rating column per participant, with a VAS column = type of rating
df_VAS <- melt(df, measure.vars= c("VAScold", "VASwarm","VASburning"), 
                        id.vars=c("ID","trial_n","manipulation","procedure",
                                  "trial_type","arm","condition","cold_location"),
             variable.name = 'VAS', value.name = 'rating')
# rename the VAS ratings
df_VAS <- df_VAS %>% 
  mutate(VAS = revalue(VAS, c("VAScold" = "cold",
                                    "VASwarm" = "warm",
                                    "VASburning" = "burning")))

# to run zero inflated regressions need to make sure no values = 100, as cannot model them, so simply minus a very small fraction from those values
df_VAS$beta <- ifelse(df_VAS$rating==100, df_VAS$beta-0.0001, df_VAS$beta <- df_VAS$rating)

# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
df_VAS$beta <- df_VAS$beta/100
df_VAS$ID <- factor(df_VAS$ID)

# Summarise data per rating, per participant
df_sum <- aggregate(rating~ID*manipulation*condition*cold_location*procedure*VAS, 
                    median, data = df_VAS)
```

# Plotting data distributions to see what kind of model family is required
```{r}
ggplot(data = df_VAS) +
  geom_density(aes(beta, fill = manipulation), alpha = .5) +
  facet_wrap(~VAS)
```

# Now to model effects! Try a zero inflated regression model
```{r}
# First model cold VAS
model.cold = glmmTMB::glmmTMB(beta ~ manipulation + condition + cold_location + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_VAS[df_VAS$VAS == 'cold' ,])
# then model warm VAS
model.warm = glmmTMB::glmmTMB(beta ~ manipulation + condition + cold_location + trial_n +
                                (1|ID),
                      family = glmmTMB::betabinomial(),
                      ziformula = ~1+manipulation,
                      data = df_VAS[df_VAS$VAS == 'warm' ,])
# model summaries
summary(model.cold)
summary(model.warm)
```
# Look at model assumptions
```{r}
# cold assumption
model.cold.assmup <- simulateResiduals(model.cold, n = 1000)
plot(model.cold.assmup)
```

```{r}
# warm assumption
model.warm.assump <- simulateResiduals(model.warm, n = 1000)
plot(model.warm.assump)
```

Cold model power analysis
# Run power analyses!!
# Cold first, then warm
# Create a model with more data points per variable, to extract more power
```{r}
# First, let's try to add 30 observations per variable combination, which would mean N = 30
model.cold.ex <- extend(model.cold, within = 'manipulation+condition+cold_location', n=30)
summary(model.cold.ex)
# check that number of data points per variable does = 30
dat_model <- xtabs(~ manipulation + condition + cold_location, data=getData(model.cold.ex))
print(dat_model)

# extract relevant effect sizes
tgi_effect <- fixef(model.cold.ex)["manipulationTGI"]
cond_effect <- fixef(model.cold.ex)["conditionwithin"]
cold_loc_effect <- fixef(model.cold.ex)["cold_locationprox_caud"]

# power analysis - using smallest effect size
fixef(model.cold.ex)["manipulationTGI"] <- cold_loc_effect
pwr.warm <- powerSim(model.warmN)
print(pwr.warm)

# power curve
# can run a power curve to determine the point of trade off between N observations and power
pc.cold <- powerCurve(model.cold.ex, within='manipulation+condition+cold_location', 
                      breaks= seq(1, 30, by = 3))
print(pc.cold)
```

# Plot power curve!
```{r}
plot(pc.cold) # a good visualisation :)
```

Now let's do the same for warm!
# Warm model
```{r}
model.warm <- lmerTest::lmer(rating ~ manipulation + condition + cold_location +
                     (1|ID),
                   data = df_sum[df_sum$VAS == 'warm' ,])
summary(model.warm)
```

# Extend the model observations to n = 50, and extract smallest effect size of interest
```{r}
# using n050 here as expecting smaller effect sizes
model.warmN <- extend(model.warm, within = 'manipulation+condition+cold_location', n=50)
# extracting effect sizes from the extended model
tgi_effect <- fixef(model.warmN)["manipulationTGI"]
cond_effect <- fixef(model.warmN)["conditionwithin"]
cold_loc_effect <- fixef(model.warmN)["cold_locationprox_caud"]
# add smallest effect size - same as cold model
fixef(model.coldN)["cold_locationprox_caud"] <- cold_loc_effect

# then run power analysis
pwr.warm <- powerSim(model.warmN)
print(pwr.warm)
```
# Run a power curve on these data with n = 50
```{r}
# can run a power curve to determine the point of trade off between N observations and power
pc.warm <- powerCurve(model.warmN, within='manipulation+condition+cold_location', 
                      breaks= seq(1, 50, by = 5))
print(pc.warm)
```