### 2022-05-17 analysis/tgispine_zoib_power-sim.Rmd
informing simulations with pilot data

---
title: "TGI-spinal-power"
author: "A.G. Mitchell"
date: '2022-04-28'
output: html_document
---

Calculating the N required for the TGI spinal experiment, by simulating data frame based on pilot data from N = 4 -- this will change!
Pre-registration link:

# Last edited: 17.05.2022

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(brms)
library(modelsummary)
library(tibble)
library(DHARMa)
library(glmmTMB)
library(reshape2)
library(truncnorm)
library(tidyverse)
library(plyr)
library(data.table)

set.seed(1234)  # Make everything reproducible

# Define the goodness-of-fit stats to include in modelsummary()
gof_stuff <- tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "N", 0,
  "r.squared", "R²", 3
)

# setting working directory
datPath <- '/Users/au706616/Documents/Experiments/SPINALTGI/Raw/'
knitr::opts_knit$set(root.dir = datPath)
```

```{r}
# import compiled data
df <- read.csv('STGI_compiled-data.csv')
```

# Reformatting data frame to extract key information
```{r}
# need to recode levels for cold_probe to reduce number of levels from 4 to 2
df$cold_location[df$cold_probe == 'distal'] <- 'dist_rostr' 
df$cold_location[df$cold_probe == 'rostral'] <- 'dist_rostr'
df$cold_location[df$cold_probe == 'proximal'] <- 'prox_caud'
df$cold_location[df$cold_probe == 'caudal'] <- 'prox_caud'
df$cold_location <- as.factor(df$cold_location)

# reshape data so that there is one rating column per participant, with a VAS column = type of rating
df_VAS <- melt(df, measure.vars= c("VAScold", "VASwarm","VASburn"), 
                        id.vars=c("ID","trial_n","manipulation","order",
                                  "trial_type","arm","condition","cold_location"),
             variable.name = 'VAS', value.name = 'rating')
# rename the VAS ratings
df_VAS <- df_VAS %>% 
  mutate(VAS = revalue(VAS, c("VAScold" = "cold",
                                    "VASwarm" = "warm",
                                    "VASburn" = "burn")))

# to run zero inflated regressions need to make sure no values = 100, as cannot model them, so simply minus a very small fraction from those values
df_VAS$beta <- ifelse(df_VAS$rating==100, df_VAS$beta-0.0001, df_VAS$beta <- df_VAS$rating)

# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
df_VAS$beta <- df_VAS$beta/100
df_VAS$ID <- factor(df_VAS$ID)

# Summarise beta per rating, per condition for all participants (for simulations)
# mean
df_sum <- aggregate(beta~manipulation*condition*cold_location*VAS, 
                    mean, data = df_VAS)
setnames(df_sum, "beta", "mean")
# standard deviation
df_sd <- aggregate(beta~manipulation*condition*cold_location*VAS, 
                    SD, data = df_VAS)
setnames(df_sd, "beta", "std")
# merge
df_sum <- merge(df_sum, df_sd)

# creating individual data-frames for specific conditions (to make simulation easier)
cnt_cold <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'cold' ,]
cnt_warm <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'warm' ,]
cnt_burn <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'burn' ,]
tgi_cold <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'cold' ,]
tgi_warm <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'warm' ,]
tgi_burn <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'burn' ,]
```

# Simulating data function
```{r}
simulate_subj <- function(subid, llim, ulim) {
# trial info
trialN = 12
conditionN = 8
ratingN = 3
# creating simulation data-frame
df <- data.frame(matrix(ncol = 0, nrow = trialN*conditionN*ratingN))
df$id <- rep(subid,trialN*conditionN*ratingN)
df$trialn <- rep(1:trialN,conditionN*ratingN)
df$dermatome <- rep(rep(c(rep("within",trialN*2), rep("across",trialN*2)),2),ratingN)
df$cold_location <- rep(rep(c(rep("dist_rostr",trialN), rep("prox_caud", trialN)),conditionN/2),ratingN)
df$stimuli <- rep(c(rep("tgi",trialN*conditionN/2), rep("non-tgi",trialN*conditionN/2)),ratingN)
df$quality <- c(rep("cold",trialN*conditionN), rep("warm",trialN*conditionN), rep("burn",trialN*conditionN))

# Simulate data here based on means and standard deviations of the above pilot sample
# Cold ratings: simulation
# tgi
w_dr_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_cold, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_cold, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_cold, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_cold, condition == 'across' & cold_location == 'prox_caud')[, 6])
# control
w_dr_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_cold, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_cold, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_cold, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_cold, condition == 'across' & cold_location == 'prox_caud')[, 6])

# Warm ratings: simulation
# tgi
w_dr_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_warm, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_warm, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_warm, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_warm, condition == 'across' & cold_location == 'prox_caud')[, 6])
# control
w_dr_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_warm, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_warm, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_warm, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_warm, condition == 'across' & cold_location == 'prox_caud')[, 6])

# Burning ratings: simulation
# tgi
w_dr_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_burn, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_burn, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_burn, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_burn, condition == 'across' & cold_location == 'prox_caud')[, 6])
# control
w_dr_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_burn, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_burn, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_burn, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_burn, condition == 'across' & cold_location == 'prox_caud')[, 6])

df$rating <- c(w_dr_tgi_cold, w_pc_tgi_cold, a_dr_tgi_cold, a_pc_tgi_cold, w_dr_nontgi_cold, w_pc_nontgi_cold, a_dr_nontgi_cold, a_pc_nontgi_cold,
               w_dr_tgi_warm, w_pc_tgi_warm, a_dr_tgi_warm, a_pc_tgi_warm, w_dr_nontgi_warm, w_pc_nontgi_warm, a_dr_nontgi_warm, a_pc_nontgi_warm,
               w_dr_tgi_burn, w_pc_tgi_burn, a_dr_tgi_burn, a_pc_tgi_burn, w_dr_nontgi_burn, w_pc_nontgi_burn, a_dr_nontgi_burn, a_pc_nontgi_burn)
return(df)}
```

# Simulate data for different N (n = 20; n = 30; n = 40)
```{r}
```

# Plotting data distributions to see what kind of model family is required
```{r}
ggplot(data = df_VAS) +
  geom_density(aes(beta, fill = manipulation), alpha = .5) +
  facet_wrap(~VAS) +
  theme_classic()
```

# Now to model effects! Try a zero inflated regression model
```{r}
# First model cold VAS
model.cold = glmmTMB::glmmTMB(beta ~ manipulation + condition + cold_location + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_VAS[df_VAS$VAS == 'cold' ,],
                      na.action = na.omit) 
# then model warm VAS
model.warm = glmmTMB::glmmTMB(beta ~ manipulation + condition + cold_location + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_VAS[df_VAS$VAS == 'warm' ,])
# model summaries
summary(model.cold)
summary(model.warm)
```
# Look at model assumptions
```{r}
# cold assumption
model.cold.assmup <- simulateResiduals(model.cold, n = 1000)
plot(model.cold.assmup)
```

```{r}
# warm assumption
model.warm.assump <- simulateResiduals(model.warm, n = 1000)
plot(model.warm.assump)
```

# Run power analyses!!
This code is extracted from [TBD]
# Cold first, then warm

```{r}
# Extract coefficient sizes
modelc.coefs

GetHyperparam<-function(x,b=NULL){
  ## Get the hyperparameters from the mixed effect model
  fe <- fixef(x)

  if(is.null(b))
    b<-fe[2] # use the data effect size if not supplied
  mu.a <- fe[1] 
  vc <- VarCorr(x)
  sigma.y <- as.numeric(vc[5, 2]) # Residual StdDev
  sigma.a <- as.numeric(vc[2, 2]) # Cobblebar StdDev
  sigma.g <- as.numeric(vc[4, 2]) # Cobblebar:transect StdDev
  hp<-c(b, mu.a, sigma.y, sigma.a, sigma.g)
  names(hp)<-c("b", "mu.a", "sigma.y", "sigma.a", "sigma.g")
  return(hp)
}
```