### 2022-05-20 analysis/tgispine_zoib_power-sim.Rmd
tidying up sim file, more logical order

---
title: "TGI-spinal-power"
author: "A.G. Mitchell"
date: '2022-04-28'
output: html_document
---

Calculating the N required for the TGI spinal experiment, by simulating data frame based on pilot data from N = 4 -- this will change!
Pre-registration link:

# Last edited: 18.05.2022

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(brms)
library(modelsummary)
library(tibble)
library(DHARMa)
library(glmmTMB)
library(reshape2)
library(truncnorm)
library(tidyverse)
library(plyr)
library(data.table)

set.seed(1234)  # Make everything reproducible

# Define the goodness-of-fit stats to include in modelsummary()
gof_stuff <- tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "N", 0,
  "r.squared", "R²", 3
)

# setting working directory
datPath <- '/home/alexm/sTGI/'
knitr::opts_knit$set(root.dir = datPath)
```

```{r}
# import compiled data
df <- read.csv('STGI_compiled-data.csv')
```

# Reformatting data frame to extract key information
```{r}
# need to recode levels for cold_probe to reduce number of levels from 4 to 2
df$cold_location[df$cold_probe == 'distal'] <- 'dist_rostr' 
df$cold_location[df$cold_probe == 'rostral'] <- 'dist_rostr'
df$cold_location[df$cold_probe == 'proximal'] <- 'prox_caud'
df$cold_location[df$cold_probe == 'caudal'] <- 'prox_caud'
df$cold_location <- as.factor(df$cold_location)

# reshape data so that there is one rating column per participant, with a VAS column = type of rating
df_VAS <- reshape2::melt(df, measure.vars= c("VAScold", "VASwarm","VASburn"),
                         id.vars=c("ID","trial_n","manipulation","order",
                                  "trial_type","arm","condition","cold_location"),
                         variable.name = 'VAS', value.name = 'rating')
# rename the VAS ratings
df_VAS <- df_VAS %>% 
  mutate(VAS = revalue(VAS, c("VAScold" = "cold",
                                    "VASwarm" = "warm",
                                    "VASburn" = "burn")))

# to run zero inflated regressions need to make sure no values = 100, as cannot model them, so simply minus a very small fraction from those values
df_VAS$beta <- ifelse(df_VAS$rating==100, df_VAS$beta-0.0001, df_VAS$beta <- df_VAS$rating)

# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
df_VAS$beta <- df_VAS$beta/100
df_VAS$ID <- factor(df_VAS$ID)

# Summarise beta per rating, per condition, per participant
# median
df_sum_pp <- aggregate(beta~ID*manipulation*condition*cold_location*VAS, 
                    median, data = df_VAS)

# now summarise across participants
df_sum <- aggregate(beta~manipulation*condition*cold_location*VAS, 
                    mean, data = df_sum_pp)
setnames(df_sum, "beta", "mean")
# standard deviation
df_sd <- aggregate(beta~manipulation*condition*cold_location*VAS, 
                    SD, data = df_sum_pp)
setnames(df_sd, "beta", "std")
# merge
df_sum <- merge(df_sum, df_sd)

# creating individual data-frames for specific conditions (to make simulation easier)
cnt_cold <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'cold' ,]
cnt_warm <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'warm' ,]
cnt_burn <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'burn' ,]
tgi_cold <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'cold' ,]
tgi_warm <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'warm' ,]
tgi_burn <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'burn' ,]
```

# Simulating data function
```{r}
simulate_subj <- function(subid, llim, ulim) {
# trial info
trialN = 12
conditionN = 8
ratingN = 3
# creating simulation data-frame
df <- data.frame(matrix(ncol = 0, nrow = trialN*conditionN*ratingN))
df$id <- rep(subid,trialN*conditionN*ratingN)
df$trialn <- rep(1:trialN,conditionN*ratingN)
df$dermatome <- rep(rep(c(rep("within",trialN*2), rep("across",trialN*2)),2),ratingN)
df$cold_location <- rep(rep(c(rep("dist_rostr",trialN), rep("prox_caud", trialN)),conditionN/2),ratingN)
df$stimuli <- rep(c(rep("tgi",trialN*conditionN/2), rep("non-tgi",trialN*conditionN/2)),ratingN)
df$quality <- c(rep("cold",trialN*conditionN), rep("warm",trialN*conditionN), rep("burn",trialN*conditionN))

# Simulate data here based on means and standard deviations of the above pilot sample
# Cold ratings: simulation
# tgi
w_dr_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_cold, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_cold, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_cold, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_tgi_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_cold, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_cold, condition == 'across' & cold_location == 'prox_caud')[, 6])
# control
w_dr_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_cold, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_cold, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_cold, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_cnt_cold <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_cold, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_cold, condition == 'across' & cold_location == 'prox_caud')[, 6])

# Warm ratings: simulation
# tgi
w_dr_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_warm, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_warm, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_warm, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_tgi_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_warm, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_warm, condition == 'across' & cold_location == 'prox_caud')[, 6])
# control
w_dr_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_warm, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_warm, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_warm, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_cnt_warm <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_warm, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_warm, condition == 'across' & cold_location == 'prox_caud')[, 6])

# Burning ratings: simulation
# tgi
w_dr_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_burn, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_burn, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(tgi_burn, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_tgi_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(tgi_burn, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(tgi_burn, condition == 'across' & cold_location == 'prox_caud')[, 6])
# control
w_dr_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'within' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_burn, condition == 'within' & cold_location == 'dist_rostr')[, 6])
w_pc_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'within' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_burn, condition == 'within' & cold_location == 'prox_caud')[, 6])
a_dr_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'across' & cold_location == 'dist_rostr')[, 5], 
    sd=filter(cnt_burn, condition == 'across' & cold_location == 'dist_rostr')[, 6])
a_pc_cnt_burn <- 
  truncnorm::rtruncnorm(
    trialN, llim, ulim, 
    mean=filter(cnt_burn, condition == 'across' & cold_location == 'prox_caud')[, 5], 
    sd=filter(cnt_burn, condition == 'across' & cold_location == 'prox_caud')[, 6])

# creating rating variable
df$rating <- c(w_dr_tgi_cold, w_pc_tgi_cold, a_dr_tgi_cold, a_pc_tgi_cold, 
               w_dr_cnt_cold, w_pc_cnt_cold, a_dr_cnt_cold, a_pc_cnt_cold,
               w_dr_tgi_warm, w_pc_tgi_warm, a_dr_tgi_warm, a_pc_tgi_warm, 
               w_dr_cnt_warm, w_pc_cnt_warm, a_dr_cnt_warm, a_pc_cnt_warm,
               w_dr_tgi_burn, w_pc_tgi_burn, a_dr_tgi_burn, a_pc_tgi_burn, 
               w_dr_cnt_burn, w_pc_cnt_burn, a_dr_cnt_burn, a_pc_cnt_burn)
return(df)}
```

# Model function - Using zero inflated regression model
```{r}
# First model cold VAS
model <- function(x, vas = c('cold','warm','burn')) {
  if (vas[1] == 'cold'){
    x1 <- x %>%  filter(quality == 'cold')
  } else if (vas[1] == 'warm'){
    x1 <- x %>%  filter(quality == 'warm')
  } else if (vas[1] == 'burn') {
    x1 <- x %>%  filter(quality == 'burn')
  }

  mod = glmmTMB::glmmTMB(
    rating ~ stimuli + dermatome + cold_location + trialn + (1|id),
    family = glmmTMB::beta_family(),
    ziformula = ~1+stimuli,
    data = x1,
    na.action = na.omit)

return(mod)}
```

# First, simulate one data-set and run models (warm, cold and burn perception) to take a look at data
```{r}
df_sim <- data.frame()
n = 50 #an example number of participants
for (i in 1:n) {
  df <- simulate_subj(i, 0, 1)
  df_sim <- rbind(df_sim,df)
}
df50 <- df_sim

# Plotting data distributions to see what kind of model family is required
df50$quality <- factor(df50$quality, 
                             levels = c("cold", "warm", "burn"))
ggplot2::ggplot(data = df50) +
  geom_density(aes(rating, fill = stimuli), alpha = .5) +
  facet_wrap(~quality) +
  theme_classic()
```

# running models based on simmed data
```{r}
# models for each VAS
mc <- model(df_sim, 'cold')
mw <- model(df_sim, 'warm')
mb <- model(df_sim, 'burn')

# show models
summary(mc)
summary(mw)
summary(mb)
```

# Look at model assumptions
```{r}
# cold assumption - on n = 50
mc.assump <- simulateResiduals(mc, n = 1000)
plot(mc.assump)
```

## SIMULATIONS ##
# Run through n = 1000 simulations of the pilot data, and extract p-value for the smallest effect (cold_location) of each model
# Extract the power from each model - first cold
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
n <- 40 # number of participants to simulate
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  df_sim <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df_sim <- rbind(df_sim,df)
    }
  # then run the model based on simulated data
  m <- model(df_sim, 'cold')
  m1 <- summary(m)
  # extracting the p-value and effect size
  tmp <- c(m1$coefficients$cond[4,4], m1$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

# rename columns
names(po)[1] <- 'pval'
names(po)[2] <- 'eff'
po$sig <- as.numeric(po$pval < .05) #is significant
po <- po[!is.na(po$sig) ,] #removing any NAs (models that did not converge) from the data frame

power <- sum(po$sig)/length(po$sig) # calculating power

# some renaming
po_c <- po
power_c <- power

power_c
```

# Then warm
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
n <- 32 # number of participants to simulate
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  df_sim <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df_sim <- rbind(df_sim,df)
    }
  # then run the model based on simulated data
  m <- model(df_sim, 'warm')
  m1 <- summary(m)
  # extracting the p-value and effect size
  tmp <- c(m1$coefficients$cond[4,4], m1$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

# rename columns
names(po)[1] <- 'pval'
names(po)[2] <- 'eff'
po$sig <- as.numeric(po$pval < .05) #is significant
po$sig[is.na(po$sig)] <- 0 #removing any NAs (code as unsig)

power <- sum(po$sig)/nSims
# some renaming
po_w <- po
power_w <- power

power_w
```

## PLOTTING - plot simulated data (d50 from earlier)
# Summary statistics for the largest data set - for plotting
```{r}
# define each condition
df50$conditionN[df50$dermatome == "within" & df50$cold_location == "dist_rostr"] <- 1
df50$conditionN[df50$dermatome == "within" & df50$cold_location == "prox_caud"] <- 2
df50$conditionN[df50$dermatome == "across" & df50$cold_location == "dist_rostr"] <- 3
df50$conditionN[df50$dermatome == "across" & df50$cold_location == "prox_caud"] <- 4

# Subject-level summary statistics
df <- df40
df_subj <- aggregate(df$rating, by = list(id = df$id, conditionN = df$conditionN, dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = mean)
#df_subj$dermatome <- factor(df_subj$dermatome, levels = c('within','across')) # re-order conditions
df_subj

# Group-level stats
df <- df_subj
group_mean   <- aggregate(df$x, by = list(conditionN = df$conditionN, dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = mean)
group_median <- aggregate(df$x, by = list(conditionN = df$conditionN,dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = median)
group_sd     <- aggregate(df$x, by = list(conditionN = df$conditionN,dermatome = df$dermatome, cold_location = df$cold_location, stimuli = df$stimuli, quality = df$quality), FUN = sd)
group_se     <- group_sd$x/sqrt(length(unique(df$id)))

df_group <- data.frame(group_mean, mean = group_mean$x, median = group_median$x, sd = group_sd$x, se = group_se)
df_group

# isolating specific data sets for plotting
df_subj_cold <- df_subj %>% filter(quality == 'cold')
df_group_cold <- df_group %>% filter(quality == 'cold')
df_subj_warm <- df_subj %>% filter(quality == 'warm')
df_group_warm <- df_group %>% filter(quality == 'warm')
df_subj_burn <- df_subj %>% filter(quality == 'burn')
df_group_burn <- df_group %>% filter(quality == 'burn')

```

## Settings for plots
```{r, echo = FALSE}
# Load libraries
library("ggplot2")
library("gghalves")
library("RColorBrewer")

# Define colors
reds <- brewer.pal(6, "Reds")
blues <- brewer.pal(6, "Blues")
purps <- brewer.pal(6, "Purples")

col_cold_n <- blues[5]
col_cold_w <- blues[3]
col_warm_n <- reds[3]
col_warm_w <- reds[5]
col_burn_n <- purps[3]
col_burn_w <- purps[5]

# Parameters for plots
tgi_theme = theme( 
    text = element_text(size = 14),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title=element_text(size=12), 
    legend.text=element_text(size=12),
    plot.title = element_text(lineheight=.8, face="bold", size = 12),
    panel.border = element_rect(size=1),
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
    legend.key = element_blank(),
    axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))

# Other parameters
nudge1 = .2
nudge2 = .3
nudge3 = .4

# Define how to plot half violins
tgi_half_violin <- function(
    df_subj, tgi_dermatome, tgi_cold_location, tgi_stimuli, nudge, color) 
  {v1 <- geom_half_violin(
    data = df_subj %>% filter(
      dermatome == tgi_dermatome & cold_location == tgi_cold_location & stimuli == tgi_stimuli), 
    aes(x = conditionN, y = x), position = position_nudge(x = nudge), 
    side = "r", fill = color, alpha = .8)
return(v1)}

# Define how to plot group means and error bars
tgi_group_mean <- function(df_group, tgi_dermatome, tgi_cold_location, tgi_stimuli, nudge, color, shape) {
m1 <- geom_point(
  data = df_group %>% filter(dermatome == tgi_dermatome & cold_location == tgi_cold_location & stimuli == tgi_stimuli), 
  aes(x = conditionN, y = mean), position = position_nudge(x = nudge + .05), size = 3, shape = shape)
return(m1)}

# Define how to plot error bars
tgi_group_error <- function(df_group, tgi_dermatome, tgi_cold_location, tgi_stimuli, tgi_nudge, tgi_color) {
e1 <- geom_errorbar(
  data = df_group %>% filter(dermatome == tgi_dermatome & cold_location == tgi_cold_location & stimuli == tgi_stimuli), 
  aes(x = conditionN, y = mean, ymin = mean - se, ymax = mean + se), position = position_nudge(tgi_nudge + .05), width = 0)
return(e1)}

```

# Plotting cold
```{r}
set.seed(321)
df_subj_cold$xj <- jitter(df_subj_cold$conditionN, amount=.1)

# cold plot
c1 <- ggplot(data = df_subj_cold, 
             mapping = aes(x = conditionN, y = x, fill = stimuli, shape = stimuli, color = stimuli)) +

  # Add theme
   theme_bw() +
   tgi_theme +
   geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5) +

   # Add points (for each subject and condition)
   geom_point(aes(x = xj), size = 2, alpha = .8) +

  # Add half violin plots
   tgi_half_violin(df_subj_cold, "within", "dist_rostr", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "within", "prox_caud", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "across", "dist_rostr", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "across", "prox_caud", "non-tgi", nudge2, col_cold_n) +
   tgi_half_violin(df_subj_cold, "within", "dist_rostr", "tgi", nudge1, col_cold_w) +
   tgi_half_violin(df_subj_cold, "within", "prox_caud", "tgi", nudge1, col_cold_w) +
   tgi_half_violin(df_subj_cold, "across", "dist_rostr", "tgi", nudge1, col_cold_w) +
   tgi_half_violin(df_subj_cold, "across", "prox_caud", "tgi", nudge1, col_cold_w) +

   # Add group summary stats (between-condition lines, errors bars, and means)
   geom_line(data = df_group_cold %>% filter(stimuli == "non-tgi"), 
             aes(x = conditionN + .45, y = mean)) +
   geom_line(data = df_group_cold %>% filter(stimuli == "tgi"), 
             aes(x = conditionN + .35, y = mean)) +

   tgi_group_error(df_group_cold, "within", "dist_rostr", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "within", "prox_caud", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "across", "dist_rostr", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "across", "prox_caud", "non-tgi", nudge3, col_cold_n) +
   tgi_group_error(df_group_cold, "within", "dist_rostr", "tgi", nudge2, col_cold_w) +
   tgi_group_error(df_group_cold, "within", "prox_caud", "tgi", nudge2, col_cold_w) +
   tgi_group_error(df_group_cold, "across", "dist_rostr", "tgi", nudge2, col_cold_w) +
   tgi_group_error(df_group_cold, "across", "prox_caud", "tgi", nudge2, col_cold_w) +

   tgi_group_mean(df_group_cold, "within", "dist_rostr", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "within", "prox_caud", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "across", "dist_rostr", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "across", "prox_caud", "non-tgi", nudge3, col_cold_n, 22) +
   tgi_group_mean(df_group_cold, "within", "dist_rostr", "tgi", nudge2, col_cold_w, 21) +
   tgi_group_mean(df_group_cold, "within", "prox_caud", "tgi", nudge2, col_cold_w, 21) +
   tgi_group_mean(df_group_cold, "across", "dist_rostr", "tgi", nudge2, col_cold_w, 21) +
   tgi_group_mean(df_group_cold, "across", "prox_caud", "tgi", nudge2, col_cold_w, 21) +

   # Define additional settings
   scale_shape_manual(values = c(22, 21)) +
   scale_x_continuous(breaks=c(1.2, 2.2, 3.2, 4.2), 
                      labels=c("distal", "proximal", "rostral","caudal"), limits=c(0.6, 5)) +
   scale_fill_manual(labels = c("non-TGI", "TGI"), values=c(blues[6],blues[3])) +
   scale_color_manual(labels = c("non-TGI", "TGI"), values=c(blues[6], "gray25")) +
   guides(shape = FALSE) +
   labs(title = "", x = "", y = "Cold ratings", fill = "Stimuli", color = "Stimuli") +
   ggtitle('Figure 1: Cold perception')

c1
```

# Plotting warm
```{r}
set.seed(321)
df_subj_warm$xj <- jitter(df_subj_warm$conditionN, amount=.1)

# cold plot
w1 <- ggplot(data = df_subj_warm, 
             mapping = aes(x = conditionN, y = x, fill = stimuli, shape = stimuli, color = stimuli)) +

  # Add theme
   theme_bw() +
   tgi_theme +
   geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5) +

   # Add points (for each subject and condition)
   geom_point(aes(x = xj), size = 2, alpha = .8) +

  # Add half violin plots
   tgi_half_violin(df_subj_warm, "within", "dist_rostr", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "within", "prox_caud", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "across", "dist_rostr", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "across", "prox_caud", "non-tgi", nudge2, col_warm_n) +
   tgi_half_violin(df_subj_warm, "within", "dist_rostr", "tgi", nudge1, col_warm_w) +
   tgi_half_violin(df_subj_warm, "within", "prox_caud", "tgi", nudge1, col_warm_w) +
   tgi_half_violin(df_subj_warm, "across", "dist_rostr", "tgi", nudge1, col_warm_w) +
   tgi_half_violin(df_subj_warm, "across", "prox_caud", "tgi", nudge1, col_warm_w) +

   # Add group summary stats (between-condition lines, errors bars, and means)
   geom_line(data = df_group_warm %>% filter(stimuli == "non-tgi"), 
             aes(x = conditionN + .45, y = mean)) +
   geom_line(data = df_group_warm %>% filter(stimuli == "tgi"), 
             aes(x = conditionN + .35, y = mean)) +

   tgi_group_error(df_group_warm, "within", "dist_rostr", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "within", "prox_caud", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "across", "dist_rostr", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "across", "prox_caud", "non-tgi", nudge3, col_warm_n) +
   tgi_group_error(df_group_warm, "within", "dist_rostr", "tgi", nudge2, col_warm_w) +
   tgi_group_error(df_group_warm, "within", "prox_caud", "tgi", nudge2, col_warm_w) +
   tgi_group_error(df_group_warm, "across", "dist_rostr", "tgi", nudge2, col_warm_w) +
   tgi_group_error(df_group_warm, "across", "prox_caud", "tgi", nudge2, col_warm_w) +

   tgi_group_mean(df_group_warm, "within", "dist_rostr", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "within", "prox_caud", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "across", "dist_rostr", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "across", "prox_caud", "non-tgi", nudge3, col_warm_n, 22) +
   tgi_group_mean(df_group_warm, "within", "dist_rostr", "tgi", nudge2, col_warm_w, 21) +
   tgi_group_mean(df_group_warm, "within", "prox_caud", "tgi", nudge2, col_warm_w, 21) +
   tgi_group_mean(df_group_warm, "across", "dist_rostr", "tgi", nudge2, col_warm_w, 21) +
   tgi_group_mean(df_group_warm, "across", "prox_caud", "tgi", nudge2, col_warm_w, 21) +

   # Define additional settings
   scale_shape_manual(values = c(22, 21)) +
   scale_x_continuous(breaks=c(1.2, 2.2, 3.2, 4.2), 
                      labels=c("distal", "proximal", "rostral","caudal"), limits=c(0.6, 5)) +
   scale_fill_manual(labels = c("non-TGI", "TGI"), values=c(col_warm_n, col_warm_w)) +
   scale_color_manual(labels = c("non-TGI", "TGI"), values=c("gray25", reds[6])) +
   guides(shape = FALSE) +
   labs(title = "", x = "", y = "Warm ratings", fill = "Stimuli", color = "Stimuli") +
   ggtitle('Figure 2: Warm perception')

w1
```
# Plotting burn
```{r}
set.seed(321)
df_subj_burn$xj <- jitter(df_subj_burn$conditionN, amount=.1)

# cold plot
b1 <- ggplot(data = df_subj_burn, 
             mapping = aes(x = conditionN, y = x, fill = stimuli, shape = stimuli, color = stimuli)) +

  # Add theme
   theme_bw() +
   tgi_theme +
   geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5) +

   # Add points (for each subject and condition)
   geom_point(aes(x = xj), size = 2, alpha = .8) +

  # Add half violin plots
   tgi_half_violin(df_subj_burn, "within", "dist_rostr", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "within", "prox_caud", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "across", "dist_rostr", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "across", "prox_caud", "non-tgi", nudge2, col_burn_n) +
   tgi_half_violin(df_subj_burn, "within", "dist_rostr", "tgi", nudge1, col_burn_w) +
   tgi_half_violin(df_subj_burn, "within", "prox_caud", "tgi", nudge1, col_burn_w) +
   tgi_half_violin(df_subj_burn, "across", "dist_rostr", "tgi", nudge1, col_burn_w) +
   tgi_half_violin(df_subj_burn, "across", "prox_caud", "tgi", nudge1, col_burn_w) +

   # Add group summary stats (between-condition lines, errors bars, and means)
   geom_line(data = df_group_burn %>% filter(stimuli == "non-tgi"), 
             aes(x = conditionN + .45, y = mean)) +
   geom_line(data = df_group_burn %>% filter(stimuli == "tgi"), 
             aes(x = conditionN + .35, y = mean)) +

   tgi_group_error(df_group_burn, "within", "dist_rostr", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "within", "prox_caud", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "across", "dist_rostr", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "across", "prox_caud", "non-tgi", nudge3, col_burn_n) +
   tgi_group_error(df_group_burn, "within", "dist_rostr", "tgi", nudge2, col_burn_w) +
   tgi_group_error(df_group_burn, "within", "prox_caud", "tgi", nudge2, col_burn_w) +
   tgi_group_error(df_group_burn, "across", "dist_rostr", "tgi", nudge2, col_burn_w) +
   tgi_group_error(df_group_burn, "across", "prox_caud", "tgi", nudge2, col_burn_w) +

   tgi_group_mean(df_group_burn, "within", "dist_rostr", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "within", "prox_caud", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "across", "dist_rostr", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "across", "prox_caud", "non-tgi", nudge3, col_burn_n, 22) +
   tgi_group_mean(df_group_burn, "within", "dist_rostr", "tgi", nudge2, col_burn_w, 21) +
   tgi_group_mean(df_group_burn, "within", "prox_caud", "tgi", nudge2, col_burn_w, 21) +
   tgi_group_mean(df_group_burn, "across", "dist_rostr", "tgi", nudge2, col_burn_w, 21) +
   tgi_group_mean(df_group_burn, "across", "prox_caud", "tgi", nudge2, col_burn_w, 21) +

   # Define additional settings
   scale_shape_manual(values = c(22, 21)) +
   scale_x_continuous(breaks=c(1.2, 2.2, 3.2, 4.2), 
                      labels=c("distal", "proximal", "rostral","caudal"), limits=c(0.6, 5)) +
   scale_fill_manual(labels = c("non-TGI", "TGI"), values=c(col_burn_n, col_burn_w)) +
   scale_color_manual(labels = c("non-TGI", "TGI"), values=c("gray25", purps[6])) +
   guides(shape = FALSE) +
   labs(title = "", x = "", y = "Burn ratings", fill = "Stimuli", color = "Stimuli") +
   ggtitle('Figure 2: Burn perception')

b1
```