### 2023-06-15 bayesVAS.Rmd
more shuffling

---
title: "Bayes VAS"
author: "A.G. Mitchell"
date: "2023-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)        # ggplot, dplyr, %>%, and friends
library(brms)             # Bayesian modeling through Stan
library(tidybayes)        # Manipulate Stan objects in a tidy way
library(broom)            # Convert model objects to data frames
library(broom.mixed)      # Convert brms model objects to data frames
library(betareg)          # Run beta regression models
library(extraDistr)       # Use extra distributions like dprop()
library(ggdist)           # Special geoms for posterior distributions
library(gghalves)         # Special half geoms
library(ggbeeswarm)       # Special distribution-shaped point jittering
library(ggrepel)          # Automatically position labels
library(patchwork)        # Combine ggplot objects
library(marginaleffects)  # Calculate marginal effects for frequentist models
library(emmeans)          # Calculate marginal effects in even fancier ways
library(modelsummary)     # Create side-by-side regression tables

set.seed(1234)  # Make everything reproducible

# Define the goodness-of-fit stats to include in modelsummary()
gof_stuff <- tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "N", 0,
  "r.squared", "R²", 3
)

# Custom ggplot theme to make pretty plots
# Get the font at https://fonts.google.com/specimen/Barlow+Semi+Condensed
theme_clean <- function() {
  theme_minimal(base_family = "Barlow Semi Condensed") +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
          axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
          strip.text = element_text(family = "BarlowSemiCondensed-Bold",
                                    size = rel(1), hjust = 0),
          strip.background = element_rect(fill = "grey80", color = NA))
}

# Make labels use Barlow by default
update_geom_defaults("label_repel", list(family = "Barlow Semi Condensed"))

anaPath <- '/Users/au706616/Documents/Experiments/SPINALTGI/Analysis/'
```

# Load own data
```{r}
setwd(anaPath)
dat <- read.csv('STGI_compiled-data.csv')

# data wrangling
dat <- dat %>% 
  select(ID, trial_n, manipulation, arm, condition, dermatome, cold_probe, order, VASburn)
head(dat)
```

```{r}
# plot data
dat$manipulation <- as.factor(dat$manipulation)

quota_halves <- ggplot(dat, aes(x = manipulation, y = VASburn)) +
  geom_half_point(aes(color = manipulation), 
                  transformation = position_quasirandom(width = 0.1),
                  side = "l", size = 0.5, alpha = 0.5) +
  geom_half_boxplot(aes(fill = manipulation), side = "r") + 
  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none", fill = "none") +
  labs(x = "Manipulation", y = "Burning VAS") +
  theme_classic()

quota_densities <- ggplot(dat, aes(x = VASburn, fill = manipulation)) +
  geom_density(alpha = 0.6) +
  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  labs(x = "Burning VAS", y = "Density", fill = "Quota") +
  theme_classic() +
  theme(legend.position = "bottom")

quota_halves | quota_densities
```
# Frequentist beta regression
```{r}
# transform VAS ratings into beta
dat1 <- dat %>% 
  mutate(VASburn = VASburn/100) %>% 
  mutate(beta = ifelse(VASburn == 0, 0.001, VASburn)) %>% 
  mutate(beta = ifelse(VASburn == 1, VASburn - 0.001, beta)) %>% 
  filter(!is.na(beta))

# run normal beta regression
model_beta <- betareg(beta ~ manipulation | manipulation, 
                      data = dat1, 
                      link = "logit")
tidy(model_beta)

# calculating model effect of manipulation
beta_mu_intercept <- model_beta %>% 
  tidy() %>% 
  filter(component == "mean", term == "(Intercept)") %>% 
  pull(estimate)

beta_mu_quota <- model_beta %>% 
  tidy() %>% 
  filter(component == "mean", term == "manipulationTGI") %>% 
  pull(estimate)

plogis(beta_mu_intercept + beta_mu_quota) - plogis(beta_mu_intercept)
```

TGI increases average burn rating by 5% - when zeros are included in the model

```{r}
beta_phi_intercept <- model_beta %>% 
  tidy() %>% 
  filter(component == "precision", term == "(Intercept)") %>% 
  pull(estimate)

beta_phi_quota <- model_beta %>% 
  tidy() %>% 
  filter(component == "precision", term == "manipulationTGI") %>% 
  pull(estimate)

no_tgi_title <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                         "), size = exp(", round(beta_phi_intercept, 2), "))")

tgi_title <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                      " + ", round(beta_mu_quota, 2), 
                      "), size = exp(", round(beta_phi_intercept, 2),
                      " + ", round(beta_phi_quota, 2), "))")

ggplot(data = tibble(x = 0:1), aes(x = x)) +
  geom_density(data = dat2, 
               aes(x = beta, fill = manipulation), alpha = 0.5, color = NA) +
  stat_function(fun = dprop, size = 1,
                args = list(size = exp(beta_phi_intercept), 
                            mean = plogis(beta_mu_intercept)),
                aes(color = no_tgi_title)) +
  stat_function(fun = dprop, size = 1,
                args = list(size = exp(beta_phi_intercept + beta_phi_quota), 
                            mean = plogis(beta_mu_intercept + beta_mu_quota)),
                aes(color = tgi_title)) +
  scale_fill_viridis_d(option = "plasma", end = 0.8, name = "Manipulation",
                       guide = guide_legend(ncol = 1, order = 1)) +
  scale_color_viridis_d(option = "plasma", end = 0.8, direction = -1, name = "",
                        guide = guide_legend(reverse = TRUE, ncol = 1, order = 2)) +
  labs(x = "Burning VAS") +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
# brms model
model_beta_bayes <- brm(
  bf(VASburn ~ manipulation,
     phi ~ quota), 
  data = dat1,
  family = (), #need to change family
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234, 
  # Use the cmdstanr backend for Stan because it's faster and more modern than
  # the default rstan You need to install the cmdstanr package first
  # (https://mc-stan.org/cmdstanr/) and then run cmdstanr::install_cmdstan() to
  # install cmdstan on your computer.
  backend = "cmdstanr"
)
```