### 2023-07-14 testing.Rmd
updates

---
title: "tester"
output: html_document
date: "2023-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, fig.width=10, fig.height=5}
experiment1 = prep_data(file.path("data", 'STGI_exp1_compiled-data.csv'))
experiment2 = prep_data(file.path("data", 'STGI_exp2_compiled-data.csv'))

df_long_exp1 = experiment1$df_long
df_long_exp2 = experiment2$df_long

######### Exp 1 (Hypothesis 1)
df_long_exp1 = df_long_exp1 %>% mutate(cold_probe = as.factor(cold_probe), manipulation = as.factor(manipulation))

##constats:

levels(df_long_exp1$cold_probe)
#within - across
within_across = c(-1/2,1/2,1/2,-1/2)
caudal_rostral = c(1,0,0,-1)
distal_proximal = c(0,1,-1,0)

cold = rbind(1/4,within_across,caudal_rostral,distal_proximal)

cold = solve(cold)

cold = cold[,-1]

model_cold_exp1 = glmmTMB::glmmTMB(beta ~ cold_probe *manipulation+trial_n+(1|ID) + (1|order),
                            family = glmmTMB::beta_family(),
                            ziformula = ~1+manipulation,
                            data = df_long_exp1 %>% filter(quality == 'cold'),
                            na.action = na.omit,
                            contrasts=list(cold_probe = cold)

model_cold_exp = glmmTMB::glmmTMB(beta ~ cold_probe *manipulation+trial_n+(1|ID) + (1|order),
                            family = glmmTMB::beta_family(),
                            ziformula = ~1+manipulation,
                            data = df_long_exp1 %>% filter(quality == 'cold'),
                            na.action = na.omit,contrasts=list(cold_probe = cold))

mm = glmmTMB::glmmTMB(beta ~ cold_probe *manipulation+trial_n+(1|ID) + (1|order),
                            family = glmmTMB::beta_family(),
                            ziformula = ~1+manipulation,
                            data = df_long_exp1 %>% filter(quality == 'cold') %>% mutate(cold_probe = relevel(cold_probe, ref = "rostral")),
                            na.action = na.omit)

summary(mm)

model_cold = glmmTMB::glmmTMB(beta ~ manipulation * condition + trial_n +
                              (1|ID) + (1|order),
                            family = glmmTMB::beta_family(),
                            ziformula = ~1+manipulation,
                            data = df_long_exp1 %>% filter(quality == 'cold'),
                            na.action = na.omit)

summary(model_cold)
summary(model_cold_exp1)
summary(model_cold_exp)

summary(model_cold_exp1)
lsmeans(model_cold_exp1, c("manipulation","cold_probe"))

summary(model_cold)
lsmeans(model_cold, c("manipulation","condition"))

model_warm = glmmTMB::glmmTMB(beta ~ manipulation * condition + trial_n +
                              (1|ID) + (1|order),
                            family = glmmTMB::beta_family(),
                            ziformula = ~1+manipulation,
                            data = df_long_exp1 %>% filter(quality == 'warm'),
                            na.action = na.omit)

plot(ggeffect(model_cold, terms = c("condition","manipulation")))

plot(ggeffect(model_warm, terms = c("condition","manipulation")))

meaneffects = df_long_exp1 %>% filter(quality == 'warm', beta != 0) %>% group_by(manipulation,condition) %>% summarize(mean = mean(beta), se = sd(beta)/sqrt(n()))

plot2 = df_long_exp1 %>% filter(quality == 'warm', beta != 0) %>% group_by(ID, manipulation,condition) %>% summarize(mean = mean(beta)) %>% ggplot(aes(x = manipulation, y = mean, col = condition))+geom_point(position = position_dodge(width = 0.3))+
  geom_pointrange(data = meaneffects, aes(x = manipulation, col = condition, y = mean, ymin = mean-2*se, ymax = mean+2*se), size = 0.5, linewidth = 1.5,position = position_dodge(width = 0.5))+theme_classic()+
  geom_line(
    data = meaneffects,
    aes(x = manipulation, y = mean, group = condition, col = condition),
    position = position_dodge(width = 0.5),
    size = 1.2)+ggtitle("Without 0")

meaneffects2 = df_long_exp1 %>% filter(quality == 'warm') %>% group_by(manipulation,condition) %>% summarize(mean = mean(beta), se = sd(beta)/sqrt(n()))

plot1 = df_long_exp1 %>% filter(quality == 'warm') %>% group_by(ID, manipulation,condition) %>% summarize(mean = mean(beta)) %>% ggplot(aes(x = manipulation, y = mean, col = condition))+geom_point(position = position_dodge(width = 0.3))+
  geom_pointrange(data = meaneffects2, aes(x = manipulation, col = condition, y = mean, ymin = mean-2*se, ymax = mean+2*se), size = 0.5, linewidth = 1.5,position = position_dodge(width = 0.5))+theme_classic()+
  geom_line(
    data = meaneffects2,
    aes(x = manipulation, y = mean, group = condition, col = condition),
    position = position_dodge(width = 0.5),
    size = 1.2)+ggtitle("With 0")

plot1+plot2

```