### 2023-08-23 testing the differencea.Rmd

---
title: "testing"
author: "jesper fischer ehmsen"
date: "2023-07-26"
output:
  html_document: default
  pdf_document: default
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(boot, broom, broom.mixed, brms, car, caret, corrplot, DHARMa, gamlss,
               ggeffects, gghalves, ggpol, ggpubr, glmmTMB, groupdata2, lme4, lmerTest, 
               modelsummary, MuMIn, polycor, RColorBrewer, rcompanion, reshape2, Rmisc,
               rsample, rstatix, simr, tibble, tidyverse, wesanderson, patchwork, cowplot,
               sjPlot
               )

source(here::here("scripts","utility.R"))
source(here::here("scripts","plots.R"))
source(here::here("scripts","models.R"))
```

Loading the data and defining the contrast:

```{r, warning=FALSE, results='hide'}
experiment2 = prep_data(file.path("data", 'STGI_exp2_compiled-data.csv'))

df_long_exp2 = experiment2$df_long %>% 
  mutate(cold_probe = as.factor(cold_probe), manipulation = as.factor(manipulation))
## Define constrats such that we both get within / across (i.e. distal & proximal vs rostral & caudual)
levels(df_long_exp2$cold_probe)
# "caudal"   "distal"   "proximal" "rostral"

#within - across
within_across = c(-1/2,1/2,1/2,-1/2)
#caudal - rostral
caudal_rostral = c(1,0,0,-1)
#distral - proximal
proximal_distal = c(0,-1,1,0)

#define the matrix
cold = rbind(1/4,within_across,caudal_rostral,proximal_distal)

#solve it 
cold = solve(cold)
```

looking at just one subject and modeling it with a normal linear regression.

First my way:

```{r, warning=FALSE}
df_long_exp1_cold = df_long_exp2 %>% filter(quality == 'cold') %>% filter(ID == 45)

model_cold_exp = lm(beta ~ manipulation * cold_probe,
                              data = df_long_exp1_cold,
                              na.action = na.omit,
                              contrasts=list(cold_probe = cold) 
                              ) 
tab_model(model_cold_exp, digits = 4)
```
<br>
<br>

The interaction between TGI and Within / across is here $\beta$ = 0.0592 p = 0.180

Then the other way:

```{r, warning=FALSE}
model_cold_exp1 = lm(beta ~ manipulation * cold_cond * condition,
                              data = df_long_exp1_cold,
                              na.action = na.omit
                              )

tab_model(model_cold_exp1, digits = 4)
```
<br>
<br>
here the interaction between TGI and Within / across is $\beta$ = -0.0456 p = 0.464

The difference seems to be the following:
<br>
if we group by the three conditions and then take only take the reference level for the cold_condition i.e. dist_rostr and then compute the difference in manipulation (control / TGI) across the difference in condition (within / across) (i.e the manipulation * condition interaction), we get the estimate from the second model i.e. -0.0456.

```{r, warning=FALSE}
df_long_exp1_cold %>% group_by(manipulation,condition, cold_cond) %>% 
  summarize(mean = mean(beta)) %>% filter(cold_cond == "dist_rostr") %>% mutate(cold_cond = NULL) %>% 
  pivot_wider(names_from = c(condition,manipulation), values_from = mean) %>%
  mutate(dif_cnt = within_CNT-across_CNT,
         dif_TGI = within_TGI-across_TGI) %>% 
  mutate(dif_dif = dif_TGI-dif_cnt) %>% .$dif_dif
```

<br>
<br>

If we on the other hand just group by manipulation and condition i.e. not filtering by dist_rostr we get the second estimate i.e. 0.0592.

```{r, warning=FALSE}
df_long_exp1_cold %>% group_by(manipulation,condition) %>% 
  summarize(mean = mean(beta)) %>% 
  pivot_wider(names_from = c(condition,manipulation), values_from = mean) %>%
  mutate(dif_cnt = within_CNT-across_CNT,dif_TGI = within_TGI-across_TGI) %>% 
  mutate(dif_dif = dif_TGI-dif_cnt) %>% .$dif_dif

```

<br>
<br>

One can also double check this with a model that only contained the manipulation and condition as factors:
```{r, warning=FALSE}
model_cold_exp2 = lm(beta ~ manipulation * condition,
                              data = df_long_exp1_cold,
                              na.action = na.omit
                              )

tab_model(model_cold_exp2, digits = 4)
```