### 2024-01-08 tgi-spinal_analysis-exp2.Rmd

---
title: "TGI Spinal Analysis Code Experiment 2"
author: "A.G. Mitchell"
date: "2022-12-02"
output: html_document
---

DATA ANALYSIS SCRIPT FOR TGI-SPINE EXPERIMENT 2

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Loading libraries
# check for pacman package and install if not found
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(boot, broom, broom.mixed, brms, car, caret, corrplot, DHARMa, gamlss,
                 ggeffects, gghalves, ggpol, ggpubr, glmmTMB, groupdata2, lme4, lmerTest, 
                 modelsummary, MuMIn, polycor, RColorBrewer, rcompanion, reshape2, Rmisc,
                 rsample, rstatix, simr, tibble, tidyverse, wesanderson
                 )
```

# Load in data
```{r paths}
# experiment 1 file
dat_file <- file.path("data", 'STGI_exp2_compiled-data.csv')
# load all data file
if (file.exists(dat_file)){
  df_res <- read.csv(dat_file)
  } else {
    print('Data file does not exist, check data is in current directory. 
          If not, run tgi-compiler.Rmd')
  }
```

# Responder check
Check whether participants median burning ratings for TGI is significantly above 0
If not, flag them as a 'non-responder'
```{r}
# Calculate median burning rating for all participants
# TGI trials only
# Then flag how many are not sig > 0
df_med <- aggregate(VASburn~ID*manipulation*cold_probe*trial_type, median, data = df_res)
tgi <-  df_med %>% 
  filter(manipulation == 'TGI')
cnt <-  df_med %>% 
  filter(manipulation == 'CNT')

# identify test results where pvalue is < .05
tgi$ID <- as.factor(tgi$ID)
i = 0
test = data.frame(matrix(nrow = 40, ncol = 3))
colnames(test) <- c('ID','pval','responder')

for (id in levels(tgi$ID)){
  i = i+1
  tmp1 <- tgi[tgi$ID == id ,]
  test$ID[i] <- id
  test$pval[i] <- (t.test(tmp1$VASburn, mu = 0, alternative = 'greater'))$p.value
  test$responder[i] <- isTRUE(test$pval[i] < .05)
}

# combine responder logic with main data-frame
test$responder <- ifelse(test$responder == TRUE, 1, 0)
test <- test[, c(1,3)]
df_res <- merge(df_res, test, by = 'ID')

# count the number of false responses
nNONRESP <- sum(test$responder == FALSE)
# print
print(paste0('Non responders: ', nNONRESP, '/', length(test$ID)))
```

# Let's look at the data!
```{r}
# Reorganise data-frame 
# pivot longer the VAS response by each quality of sensation
# do this for both RT and VAS
df_long <- df_res %>% 
  pivot_longer(cols = c(VASinit, VASburn, VASwarm, VAScold),
               names_to = 'quality',
               values_to = 'VAS')
df_long2 <- df_res %>% 
  pivot_longer(cols = c(RTinit, RTburn, RTwarm, RTcold),
               names_to = 'quality',
               values_to = 'RT')

# remove the VAS & RT from quality column
df_long$quality <- substr(df_long$quality, 4, 7)
df_long2$quality <- substr(df_long2$quality, 3, 6)

# remove columns from both then merge
df_long <- df_long %>% 
  select(-c(coolTemp, warmTemp, Order, Manipulation.first))
df_long2 <- df_long2 %>% 
  select(-c(coolTemp, warmTemp, Order, Manipulation.first))

df_long <- merge(df_long, df_long2)

# then remove NaNs
df_long <- df_long %>% filter(!is.na(VAS))
# remove the initial trials - not important for now
init_burn <- filter(df_long, quality == 'init')
all_vas <- filter(df_long, quality != 'init')

# plot distribution of data
ggplot(df_long) +
  geom_density(aes(VAS, colour = manipulation)) +
  facet_wrap(~quality, scales = "free_y") +
  theme_bw()

# summary statistics
```
# MAIN HYPOTHESES
# Plot descriptives for each hypotheses
Hypothesis 1: within vs. across dermatomes

```{r h1}
# calculate means
vas_meds <- aggregate(VAS ~ quality*manipulation*condition*cold_probe*ID, 
                      median, data = all_vas)
vas_h1 <- aggregate(VAS ~ quality*manipulation*condition*ID, mean, data = vas_meds)
# change name of manipulation
vas_h1$manipulation <- factor(vas_h1$manipulation,
                           labels = c('Non-TGI', 'TGI'))
# summary means for h1
h1_sum <- summarySEwithin(data = vas_h1, measurevar = 'VAS',
                               withinvars = c('manipulation', 'quality', 'condition'))

# plot the summary statistics
# first colours
blue <- brewer.pal(8, "Blues")
grey <- brewer.pal(8, "Greys")
purp <- brewer.pal(8, "Purples")
oran <- brewer.pal(8, "Oranges")
gren <- brewer.pal(8, "Greens")

# order conditions
h1_sum$condition <- factor(h1_sum$condition, levels = c('within', 'across'))
h1_sum$quality <- factor(h1_sum$quality, levels = c('cold', 'warm', 'burn'))

# plot
ggplot(h1_sum, aes(colour = quality, group = manipulation, alpha = manipulation)) + 
  geom_point(aes(condition, VAS), size = 2, position = position_dodge(-.3)) +
  geom_errorbar(aes(condition, ymin = VAS-ci, ymax = VAS+ci), width = .1, size = .7,
                position = position_dodge(-.3)) +
  geom_line(aes(condition, VAS), size = .7, position = position_dodge(-.3)) +
  facet_wrap(~quality) +
  scale_colour_manual(values = c(blue[5], oran[5], purp[5])) +
  scale_alpha_discrete(range = c(0.6, 1)) +
  ylim(0,50) +
  theme_bw() 
```

# Hypothesis 1: plot difference across conditions of interest
```{r h1diff}
# then pivot wider and calculate difference
vas_h1_diff <- vas_h1 %>% 
  pivot_wider(names_from = condition, values_from = VAS) %>% 
  mutate(difference = within - across)
# recode quality
vas_h1_diff$quality <- factor(vas_h1_diff$quality,
                           levels = c('cold', 'warm', 'burn'))

# summary statistics
h1_diff_sum <- summarySEwithin(data = vas_h1_diff, measurevar = 'difference',
                               withinvars = c('manipulation', 'quality'))

# now plot the difference
ggplot(data = vas_h1_diff, aes(colour = quality)) +
  geom_hline(yintercept = 0, colour = 'grey50') +
  geom_point(aes(manipulation, difference, group = ID), position = position_dodge(.2),
             alpha = .5) +
  geom_point(data = h1_diff_sum, aes(manipulation, difference), colour = 'grey15',
             fill = 'grey15',
             shape = 21, size = 3) +
  geom_errorbar(data = h1_diff_sum, aes(manipulation, ymin = difference-ci,
                                        ymax = difference+ci), 
                width = .1, size = .7, colour = 'grey15') +
  facet_wrap(~quality) +
  scale_colour_manual(values = c(blue[5], oran[5], purp[5])) +
  labs(title = 'Hypothesis 1: Segmental Distance',
       y = 'Within - Across VAS Ratings', x = NULL) +
  theme_classic() +
  theme(legend.position = 'none') -> h1_diff_plot

h1_diff_plot

ggsave(here::here("figures", "E2_H1_difference.png"), plot = h1_diff_plot, 
       device = NULL, width = 9, height = 5, dpi = 600)
```

Hypothesis 2 a & b: relative location of warm and cold
````{r h2}
# Change name of manipulation
vas_meds$manipulation <- factor(vas_meds$manipulation, labels = c('Non-TGI', 'TGI'))
h2_sum <- aggregate(VAS~quality*manipulation*cold_probe, median, data = vas_meds)

# get confidence intervals
h2_conf <- aggregate(VAS~quality*manipulation*cold_probe, IQR, data = vas_meds)
h2_conf <- h2_conf %>% 
  dplyr::rename(iqr = VAS)

h2_sum <- merge(h2_sum, h2_conf, by = c('quality','manipulation','cold_probe'))

# organise data for plotting
vas_meds$cold_probe <- factor(vas_meds$cold_probe, levels = 
                                c('proximal', 'distal', 'rostral', 'caudal'))
h2_sum$cold_probe <- factor(h2_sum$cold_probe, labels = 
                                c('proximal', 'distal', 'rostral', 'caudal'))
# recoding cold probe location so can jitter
vas_meds$cold_code <- factor(vas_meds$cold_probe, labels = 
                                c(1,2,3,4))
h2_sum$cold_code <- factor(h2_sum$cold_probe, labels = 
                                c(1,2,3,4))

```

# Plot all summary data - for poster :)
```{r}
ggplot(data = h2_sum, aes(group = manipulation, colour = manipulation)) +
  geom_point(aes(cold_probe, VAS),
             position = position_dodge(.3), size = 5) +
  geom_errorbar(aes(x = cold_probe, ymin = VAS-iqr, ymax = VAS+iqr),
                position = position_dodge(.3), width = .2, linewidth = 1.1) +
  geom_line(aes(cold_probe, VAS), position = position_dodge(.3),
            linewidth = 1.7) +
  labs(y = 'Burn ratings (0 - 100)', x = '') +
  scale_colour_manual(values = c(purp[5], purp[8])) +
  #ylim(, 60) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)) -> h2_means

# save
ggsave(here::here("figures", "E2_H2_means.png"), plot = h2_means, 
       device = NULL, width = 7, height = 5, dpi = 600)
```

# All dat plot
```{r}
# creating jitter
vas_meds$xj <- jitter(as.numeric(vas_meds$cold_code), amount = .1)

nudge1 = -.15
nudge2 = .15

# plot all data
# COLD
ggplot(data = vas_meds %>% 
         filter(quality == 'cold'),
       mapping = aes(x = xj, y = VAS, colour = manipulation, fill = manipulation)) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'proximal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'proximal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'distal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'distal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'rostral', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'rostral', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'caudal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'caudal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_line(data = h2_sum %>% 
    filter(quality == 'cold', manipulation == 'Non-TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge1)) + 
  geom_line(data = h2_sum %>% 
    filter(quality == 'cold', manipulation == 'TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge2)) +
  scale_colour_manual(values = c(blue[5], blue[7])) +
  scale_fill_manual(values = c(blue[2], blue[4])) +
  scale_x_continuous(breaks = c(1,2,3,4), labels = c('proximal','distal','rostral','caudal')) +
  labs(title = 'Cold ratings', x = NULL, 
         y = 'VAS Rating (0-100)') +
  ylim(-0.5,100) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12),
        title = element_text(size = 12)) -> h2_cold_plot

# WARM
ggplot(data = vas_meds %>% 
         filter(quality == 'warm'),
       mapping = aes(x = xj, y = VAS, colour = manipulation, fill = manipulation)) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'proximal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'proximal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'distal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'distal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'rostral', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'rostral', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'caudal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'caudal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_line(data = h2_sum %>% 
    filter(quality == 'warm', manipulation == 'Non-TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge1)) + 
  geom_line(data = h2_sum %>% 
    filter(quality == 'warm', manipulation == 'TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge2)) +
  scale_colour_manual(values = c(oran[5], oran[7])) +
  scale_fill_manual(values = c(oran[2], oran[4])) +
  scale_x_continuous(breaks = c(1,2,3,4), labels = c('proximal','distal','rostral','caudal')) +
  labs(title = 'Warm ratings', x = NULL, 
         y = 'VAS Rating (0-100)') +
  ylim(-0.5,100) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12),
        title = element_text(size = 12)) -> h2_warm_plot

# BURN
ggplot(data = vas_meds %>% 
         filter(quality == 'burn'),
       mapping = aes(x = xj, y = VAS, colour = manipulation, fill = manipulation)) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'proximal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'proximal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'distal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'distal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'rostral', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'rostral', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'caudal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_boxjitter(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'caudal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), width = .2,
      errorbar.length = .2, jitter.shape = 21, jitter.size = 1.5,
      outlier.shape = NA, errorbar.draw = TRUE, lwd = 0.7) +
  geom_line(data = h2_sum %>% 
    filter(quality == 'burn', manipulation == 'Non-TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge1)) + 
  geom_line(data = h2_sum %>% 
    filter(quality == 'burn', manipulation == 'TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge2)) +
  scale_colour_manual(values = c(purp[5], purp[7])) +
  scale_fill_manual(values = c(purp[2], purp[4])) +
  scale_x_continuous(breaks = c(1,2,3,4), labels = c('proximal','distal','rostral','caudal')) +
  labs(title = 'Burn ratings', x = NULL, 
         y = 'VAS Rating (0-100)') +
  ylim(-0.5,100) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12),
        title = element_text(size = 12))  -> h2_burn_plot

# arrange entire plot
h2_plot <- ggarrange(h2_cold_plot, h2_warm_plot, h2_burn_plot,
                     nrow = 2, ncol = 2)
h2_plot

ggsave(here::here("figures", "E2_H2_allDat.png"), plot = h2_plot, 
       device = NULL, width = 9, height = 5, dpi = 600)
```

# make box plots for poster
```{r}
# cold 
ggplot(data = vas_meds %>% filter(quality == 'cold'),
        mapping = aes(x = xj, y = VAS, colour = manipulation, fill = manipulation)) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'proximal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'proximal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'distal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'distal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'rostral', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'rostral', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'caudal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'cold', cold_probe == 'caudal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_line(data = h2_sum %>% 
    filter(quality == 'cold', manipulation == 'Non-TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge1), linewidth = 1.5) + 
  geom_line(data = h2_sum %>% 
    filter(quality == 'cold', manipulation == 'TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge2), linewidth = 1.5) +
  scale_colour_manual(values = c(blue[5], blue[7])) +
  scale_fill_manual(values = c(blue[2], blue[4])) +
  scale_x_continuous(breaks = c(1,2,3,4), labels = c('proximal','distal','rostral','caudal')) +
  labs(x = NULL, y = '') +
  ylim(-0.5,100) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 19)) -> cold_box

# warm
ggplot(data = vas_meds %>% filter(quality == 'warm'),
        mapping = aes(x = xj, y = VAS, colour = manipulation, fill = manipulation)) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'proximal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'proximal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'distal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'distal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'rostral', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'rostral', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'caudal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'warm', cold_probe == 'caudal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_line(data = h2_sum %>% 
    filter(quality == 'warm', manipulation == 'Non-TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge1), linewidth = 1.5) + 
  geom_line(data = h2_sum %>% 
    filter(quality == 'warm', manipulation == 'TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge2), linewidth = 1.5) +
  scale_colour_manual(values = c(oran[5], oran[7])) +
  scale_fill_manual(values = c(oran[2], oran[4])) +
  scale_x_continuous(breaks = c(1,2,3,4), labels = c('proximal','distal','rostral','caudal')) +
  labs(x = NULL, y = '') +
  ylim(-0.5,100) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 19)) -> warm_box

# burn
ggplot(data = vas_meds %>% filter(quality == 'burn'),
        mapping = aes(x = xj, y = VAS, colour = manipulation, fill = manipulation)) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'proximal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'proximal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'distal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'distal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'rostral', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'rostral', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'caudal', manipulation == 'Non-TGI'),
    position = position_nudge(x = nudge1), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_boxplot(data = vas_meds %>% 
    filter(quality == 'burn', cold_probe == 'caudal', manipulation == 'TGI'),
    position = position_nudge(x = nudge2), size = 1.2, outlier.size = 2, 
    outlier.alpha = .5) +
  geom_line(data = h2_sum %>% 
    filter(quality == 'burn', manipulation == 'Non-TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge1), linewidth = 1.5) + 
  geom_line(data = h2_sum %>% 
    filter(quality == 'burn', manipulation == 'TGI'),
    aes(as.numeric(cold_code), VAS), 
    position = position_nudge(nudge2), linewidth = 1.5) +
  scale_colour_manual(values = c(purp[5], purp[7])) +
  scale_fill_manual(values = c(purp[2], purp[4])) +
  scale_x_continuous(breaks = c(1,2,3,4), labels = c('proximal','distal','rostral','caudal')) +
  labs(x = NULL, y = '') +
  ylim(-0.5,100) +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 19)) -> burn_box
```

# put them all together
```{r}
p_plot <- ggarrange(cold_box, warm_box, burn_box,
                     nrow = 3, ncol = 1)
p_plot

ggsave(here::here("figures", "E2_posterPlot.png"), plot = p_plot, 
       device = NULL, width = 7, height = 15, dpi = 600)
```

# Differences
```{r h2diff}
# recode cold_location and condition to reduce levels in regression
vas_meds$cold_cond[vas_meds$cold_probe == 'distal'] <- 'dist_rost' 
vas_meds$cold_cond[vas_meds$cold_probe == 'rostral'] <- 'dist_rost' 
vas_meds$cold_cond[vas_meds$cold_probe == 'proximal'] <- 'prox_caud' 
vas_meds$cold_cond[vas_meds$cold_probe == 'caudal'] <- 'prox_caud'

# then pivot wider and calculate difference
vas_h2_diff <- vas_meds %>% 
  select(-c(cold_code, xj, cold_probe)) %>% 
  pivot_wider(id_cols = c(ID, quality, manipulation, condition), 
              names_from = cold_cond, values_from = VAS) %>% 
  mutate(difference = prox_caud - dist_rost)
# recode quality
vas_h2_diff$quality <- factor(vas_h2_diff$quality,
                           levels = c('cold', 'warm', 'burn'))
# summary statistics
h2_diff_sum <- summarySEwithin(data = vas_h2_diff, measurevar = 'difference',
                               withinvars = c('manipulation', 'quality', 'condition'))

# plot the differences - within
ggplot(data = vas_h2_diff %>% 
         filter(condition == 'within'), 
       aes(manipulation, difference, colour = quality)) +
  geom_hline(yintercept = 0) +
  geom_point(aes(group = ID), position = position_dodge(.2), alpha = .5) +
  geom_point(data = h2_diff_sum %>% 
               filter(condition == 'within'), 
             aes(manipulation, difference), colour = 'grey15',
             fill = 'grey15',
             shape = 21, size = 3) +
  geom_errorbar(data = h2_diff_sum %>% 
               filter(condition == 'within'), 
               aes(manipulation, ymin = difference-ci, ymax = difference+ci),
               width = .1, size = .7, colour = 'grey15') +
  scale_colour_manual(values = c(blue[5], oran[5], purp[5])) +
  labs(title = 'Within: Proximal vs. Distal',
       y = 'Proximal - Distal VAS Ratings', x = NULL) +
  facet_wrap(~quality) +
  theme_classic() +
  theme(legend.position = 'none') -> h2_diff_within

# plot the differences - across
ggplot(data = vas_h2_diff %>% 
         filter(condition == 'across'), 
       aes(manipulation, difference, colour = quality)) +
  geom_hline(yintercept = 0) +
  geom_point(aes(group = ID), position = position_dodge(.2), alpha = .5) +
  geom_point(data = h2_diff_sum %>% 
               filter(condition == 'across'), 
             aes(manipulation, difference), colour = 'grey15',
             fill = 'grey15',
             shape = 21, size = 3) +
  geom_errorbar(data = h2_diff_sum %>% 
               filter(condition == 'across'), 
               aes(manipulation, ymin = difference-ci, ymax = difference+ci),
               width = .1, size = .7, colour = 'grey15') +
  scale_colour_manual(values = c(blue[5], oran[5], purp[5])) +
  labs(title = 'Across: Rostral vs. Caudal',
       y = 'Caudal - Rostral VAS Ratings', x = NULL) +
  facet_wrap(~quality) +
  theme_classic() +
  theme(legend.position = 'none') -> h2_diff_across

# arrange entire plot
h2diff_plot <- ggarrange(h2_diff_within, h2_diff_across,
                     nrow = 2, ncol = 1)
h2diff_plot

ggsave(here::here("figures", "E2_H2_difference.png"), plot = h2diff_plot, 
       device = NULL, width = 9, height = 5, dpi = 600)
```

# Plot 32 burn responders seperately
```{r burn32}
```

# ZOIB regression
Running main analysis of a zero inflated beta regression
Outcome: VAS
Predictors: manipulation (CNT, TGI), cold probe location ()
Run a seperate regression for each quality

```{r model_prep}
set.seed(1234)  # Make everything reproducible

# Define the goodness-of-fit stats to include in modelsummary()
gof_stuff <- tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "N", 0,
  "r.squared", "R²", 3
)

# recode cold_location and condition to reduce levels in regression
df_long$cold_cond[df_long$cold_probe == 'distal'] <- 'dist_rostr' 
df_long$cold_cond[df_long$cold_probe == 'rostral'] <- 'dist_rostr' 
df_long$cold_cond[df_long$cold_probe == 'proximal'] <- 'prox_caud' 
df_long$cold_cond[df_long$cold_probe == 'caudal'] <- 'prox_caud'

# to run zero inflated regressions need to make sure no values = 100, 
# as cannot model them, so simply minus a very small fraction from those values
df_long$beta <- ifelse(df_long$VAS==100, df_long$beta-0.0001, df_long$beta <- df_long$VAS)
df_long$beta <- df_long$beta/100

# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
#df_long$VAS <- df_long$VAS/100
df_long$ID <- factor(df_long$ID)

df_long_exp2 = df_long
```

Running models as pre-registered: 
# Cold
```{r coldmod}
# Try and model all with ratings
model.cold = glmmTMB::glmmTMB(beta ~ manipulation * condition * cold_cond + trial_n +
                                (1|ID) + (1|order),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_long_exp2 %>% filter(quality == 'cold'),
                      na.action = na.omit) 
mc <- summary(model.cold)
mc

Anova(model.cold)
# get Rsq with the mumin package
```

# Marginal effect plots of full model - cold
```{r}
cold_eff <- ggeffects::ggpredict(model.cold, terms = c('cold_cond', 'manipulation', 'condition'),
                           ci.lvl = .95)
# changing labels for plotting
cold_eff$group <- factor(cold_eff$group, labels = c('Non-TGI', 'TGI'))
cold_eff <- cold_eff %>% 
  mutate(label = 
           case_when(facet == 'across' & x == 'prox_caud' ~ 'caudal',
                     facet == 'across' & x == 'dist_rostr' ~ 'rostral',
                     facet == 'within' & x == 'prox_caud' ~ 'proximal',
                     facet == 'within' & x == 'dist_rostr' ~ 'distal',) 
         )

cold_eff$label = factor(cold_eff$label, levels = c('proximal','distal','caudal','rostral'))

# plotting - within
ggplot(data = cold_eff %>% 
         filter(facet == 'within'), 
       aes(label, predicted, group = group, colour = group),) +
  geom_point(position = position_dodge(.3), size = 2.5) +
  geom_line(position = position_dodge(.3), size = 1) +
  geom_errorbar(aes(label, ymin = conf.low, ymax = conf.high),
                position = position_dodge(.3), width = .1, size = .8) +
  ylim(0, 0.5) +
  labs(x = NULL, y = 'Predicted Cold Rating (0 - 1)') +
  facet_wrap(~facet) +
  scale_colour_manual(values = c(blue[4], blue[7])) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.text = element_text(size = 11)) -> plot_coldw_eff
# plotting - across
ggplot(data = cold_eff %>% 
         filter(facet == 'across'), 
       aes(label, predicted, group = group, colour = group),) +
  geom_point(position = position_dodge(.3), size = 2.5) +
  geom_line(position = position_dodge(.3), size = 1) +
  geom_errorbar(aes(label, ymin = conf.low, ymax = conf.high),
                position = position_dodge(.3), width = .1, size = .8) +
  ylim(0, 0.5) +
  labs(x = NULL, y = '') +
  facet_wrap(~facet) +
  scale_colour_manual(values = c(blue[4], blue[7])) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.text = element_text(size = 10)) -> plot_colda_eff

# combine
plot_cold_eff1 <- ggarrange(plot_coldw_eff, plot_colda_eff,
                           ncol = 2, nrow = 1,
                           common.legend = TRUE,
                           legend = 'right')
plot_cold_eff1
```

# Warm
```{r warmmod}
# Try and model all with ratings
model.warm = glmmTMB::glmmTMB(beta ~ manipulation * condition * cold_cond +
                                trial_n + (1|ID) + (1|order),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_long_exp2 %>% filter(quality == 'warm'),
                      na.action = na.omit) 
mw <- summary(model.warm)
mw

Anova(model.warm)
# save results
```

# Marginal effect plots of full model - warm
```{r}
warm_eff <- ggeffects::ggpredict(model.warm, terms = c('cold_cond', 'manipulation', 'condition'),
                            ci.lvl = .95)
# changing labels for plotting
warm_eff$group <- factor(warm_eff$group, labels = c('Non-TGI', 'TGI'))
warm_eff <- warm_eff %>% 
  mutate(label = 
           case_when(facet == 'across' & x == 'prox_caud' ~ 'caudal',
                     facet == 'across' & x == 'dist_rostr' ~ 'rostral',
                     facet == 'within' & x == 'prox_caud' ~ 'proximal',
                     facet == 'within' & x == 'dist_rostr' ~ 'distal',) 
         )

warm_eff$label = factor(warm_eff$label, levels = c('proximal','distal','caudal','rostral'))

# plotting - within
ggplot(data = warm_eff %>% 
         filter(facet == 'within'), 
       aes(label, predicted, group = group, colour = group),) +
  geom_point(position = position_dodge(.3), size = 2.5) +
  geom_line(position = position_dodge(.3), size = 1) +
  geom_errorbar(aes(label, ymin = conf.low, ymax = conf.high),
                position = position_dodge(.3), width = .1, size = .8) +
  ylim(0, 0.5) +
  labs(x = NULL, y = 'Predicted Warm Rating (0 - 1)') +
  facet_wrap(~facet) +
  scale_colour_manual(values = c(oran[4], oran[7])) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.text = element_text(size = 11)) -> plot_warmw_eff
# plotting - across
ggplot(data = warm_eff %>% 
         filter(facet == 'across'), 
       aes(label, predicted, group = group, colour = group),) +
  geom_point(position = position_dodge(.3), size = 2.5) +
  geom_line(position = position_dodge(.3), size = 1) +
  geom_errorbar(aes(label, ymin = conf.low, ymax = conf.high),
                position = position_dodge(.3), width = .1, size = .8) +
  ylim(0, 0.5) +
  labs(x = NULL, y = '') +
  facet_wrap(~facet) +
  scale_colour_manual(values = c(oran[4], oran[7])) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.text = element_text(size = 10)) -> plot_warma_eff

# combine
plot_warm_eff1 <- ggarrange(plot_warmw_eff, plot_warma_eff,
                           ncol = 2, nrow = 1,
                           common.legend = TRUE,
                           legend = 'right')
plot_warm_eff1
```
# Burn
The burning hypothesis is only for participants that experience burning TGI 
(aka are responders)
```{r burnmod}
# First, remove non-responders
df_resp_exp2 <- df_long_exp2 %>% 
  filter(responder == 1)
# check n
n <- count(df_resp_exp2, 'ID')
length(n$ID)
# Try and model all with ratings
model.burn = glmmTMB::glmmTMB(beta ~ manipulation * condition * cold_cond + 
                                trial_n + (1|ID) + (1|order),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_resp_exp2 %>% filter(quality == 'burn'),
                      na.action = na.omit) 
mb <- summary(model.burn)
mb

Anova(model.burn)
```

# Marginal effect plots of full model - burn
```{r}
burn_eff <- ggeffects::ggpredict(model.burn, terms = c('cold_cond', 'manipulation', 'condition'),
                            ci.lvl = .95)
# changing labels for plotting
burn_eff$group <- factor(burn_eff$group, labels = c('Non-TGI', 'TGI'))
burn_eff <- burn_eff %>% 
  mutate(label = 
           case_when(facet == 'across' & x == 'prox_caud' ~ 'caudal',
                     facet == 'across' & x == 'dist_rostr' ~ 'rostral',
                     facet == 'within' & x == 'prox_caud' ~ 'proximal',
                     facet == 'within' & x == 'dist_rostr' ~ 'distal',) 
         )

burn_eff$label = factor(burn_eff$label, levels = c('proximal','distal','caudal','rostral'))

# plotting - within
ggplot(data = burn_eff %>% 
         filter(facet == 'within'), 
       aes(label, predicted, group = group, colour = group),) +
  geom_point(position = position_dodge(.3), size = 2.5) +
  geom_line(position = position_dodge(.3), size = 1) +
  geom_errorbar(aes(label, ymin = conf.low, ymax = conf.high),
                position = position_dodge(.3), width = .1, size = .8) +
  ylim(0, 0.5) +
  labs(x = NULL, y = 'Predicted Burn Rating (0 - 1)') +
  facet_wrap(~facet) +
  scale_colour_manual(values = c(purp[4], purp[7])) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.text = element_text(size = 11)) -> plot_burnw_eff
# plotting - across
ggplot(data = burn_eff %>% 
         filter(facet == 'across'), 
       aes(label, predicted, group = group, colour = group),) +
  geom_point(position = position_dodge(.3), size = 2.5) +
  geom_line(position = position_dodge(.3), size = 1) +
  geom_errorbar(aes(label, ymin = conf.low, ymax = conf.high),
                position = position_dodge(.3), width = .1, size = .8) +
  ylim(0, 0.5) +
  labs(x = NULL, y = '') +
  facet_wrap(~facet) +
  scale_colour_manual(values = c(purp[4], purp[7])) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.text = element_text(size = 10)) -> plot_burna_eff

# combine
plot_burn_eff1 <- ggarrange(plot_burnw_eff, plot_burna_eff,
                           ncol = 2, nrow = 1,
                           common.legend = TRUE,
                           legend = 'right')
plot_burn_eff1
```

# Zero inflation marginal effects plot of all models
```{r zero_plots}
# Getting the marginal effects for plotting
# First the zeros - create new lm mo
df_long$bin <- ifelse(df_long$VAS == 0, 1, 0)
df_resp$bin <- ifelse(df_resp$VAS == 0, 1, 0)

# model for each VAS
# cold
cold.zeff <- glm(bin~1+manipulation, 
                  family = binomial(link = 'logit'),
                  data = df_long %>% filter(quality == 'cold'))
summary(cold.zeff)
# warm
warm.zeff <- glm(bin~1+manipulation, 
                  family = binomial(link = 'logit'),
                  data = df_long %>% filter(quality == 'warm'))
summary(warm.zeff)
# burn
burn.zeff <- glm(bin~1+manipulation, 
                  family = binomial(link = 'logit'),
                  data = df_resp %>% filter(quality == 'burn'))
summary(burn.zeff)

# Data-frames
cold_zeff <- ggeffects::ggeffect(cold.zeff, terms = c('manipulation'))
warm_zeff <- ggeffects::ggeffect(warm.zeff, terms = c('manipulation'))
burn_zeff <- ggeffects::ggeffect(burn.zeff, terms = c('manipulation'))
# changing labels for plotting
cold_zeff$x <- factor(cold_zeff$x, labels = c('Non-TGI', 'TGI'))
warm_zeff$x <- factor(warm_zeff$x, labels = c('Non-TGI', 'TGI'))
burn_zeff$x <- factor(burn_zeff$x, labels = c('Non-TGI', 'TGI'))

# plotting - cold
ggplot(cold_zeff, aes(x, predicted, colour = x)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(x, ymin = conf.low, ymax = conf.high),
                width = .05, size = .8) +
  labs(y = 'Predicted proportion of zeros', x = NULL) +
  ylim(0,.25) +
  scale_colour_manual(values = c(blue[4], blue[7])) +
  theme_classic() + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) -> plot_zcold

# warm
ggplot(warm_zeff, aes(x, predicted, colour = x)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(x, ymin = conf.low, ymax = conf.high),
                width = .05, size = .8) +
  labs(y = 'Predicted proportion of zeros', x = NULL) +
  ylim(.45,.7) +
  scale_colour_manual(values = c(oran[4], oran[7])) +
  theme_classic() + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) -> plot_zwarm

# burn
ggplot(burn_zeff, aes(x, predicted, colour = x)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(x, ymin = conf.low, ymax = conf.high),
                width = .05, size = .8) +
  labs(y = 'Predicted proportion of zeros', x = NULL) +
  ylim(.05,.3) +
  scale_colour_manual(values = c(purp[4], purp[7])) +
  theme_classic() + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) -> plot_zburn
```

# Compile all marginal effect plots
```{r}
# Cold
plot_cold_eff <- ggarrange(plot_zcold, plot_cold_eff1,
                   ncol = 2, nrow = 1,
                   widths = c(.35,1))
# add title
plot_cold_eff <- annotate_figure(plot_cold_eff, top = text_grob("Cold Ratings", 
                                                size = 12, hjust = 4.05))
plot_cold_eff

# Warm
plot_warm_eff <- ggarrange(plot_zwarm, plot_warm_eff1,
                   ncol = 2, nrow = 1,
                   widths = c(.35,1))
# add title
plot_warm_eff <- annotate_figure(plot_warm_eff, top = text_grob("Warm Ratings", 
                                                size = 12, hjust = 3.69))
plot_warm_eff

# Burn
plot_burn_eff <- ggarrange(plot_zburn, plot_burn_eff1,
                   ncol = 2, nrow = 1,
                   widths = c(.35,1))
# add title
plot_burn_eff <- annotate_figure(plot_burn_eff, top = text_grob("Burn Ratings", 
                                                size = 12, hjust = 4.1))
plot_burn_eff

# Compile all!
plot_eff <- ggarrange(plot_cold_eff, plot_warm_eff, plot_burn_eff,
                      ncol = 1, nrow = 3)
plot_eff

ggsave(here::here("figures", "E2_marginalEffects.png"), plot = plot_eff, 
       device = NULL, width = 9, height = 5, dpi = 600)
```

# Run model checks on all
```{r checks, warning=FALSE}
# cold assumption - on n = 50
mc.assump <- simulateResiduals(model.cold, n = 1000)
plot(mc.assump)
# warm assumption - on n = 50
mw.assump <- simulateResiduals(model.warm, n = 1000)
plot(mw.assump)
# warm assumption - on n = 50
mb.assump <- simulateResiduals(model.burn, n = 1000)
plot(mb.assump)
```

# EXPLORATORY HYPOTHESES
# Correlation between quality (cold/warm/burn)

```{r corr}
# remove columns
df_corr <- df_res[, c(1:9,14:16,21)]
df_corr <- df_corr %>% 
  pivot_wider(names_from = 'manipulation', values_from = c('VAScold','VASwarm','VASburn')) %>% 
  select(c(ID,VAScold_CNT,VAScold_TGI,VASwarm_CNT,VASwarm_TGI,VASburn_CNT,VASburn_TGI)) %>% 
  rename(CNTcold = VAScold_CNT,
         TGIcold = VAScold_TGI,
         CNTwarm = VASwarm_CNT,
         TGIwarm = VASwarm_TGI,
         CNTburn = VASburn_CNT,
         TGIburn = VASburn_TGI)

# average values within participants
df_corr <- aggregate(. ~ID, mean, data = df_corr)
# remove ID
df_corr <- select(df_corr, -ID)

# check distribution of data
df_means <- aggregate(VAS ~ ID*quality*manipulation, mean, data = vas_meds)
ggplot(df_means, aes(colour = quality)) +
  geom_density(aes(VAS)) +
  theme_bw()

# correlate
qual_cor <- hetcor(df_corr, type = 'spearman')
cormat <- qual_cor$correlations

# significance values
p <- cor.mtest(df_corr, type ='spearman')$p

# correlation matrix
# plotting correlations between variables
corrplot(cormat, type = 'lower',
         p.mat = p, insig = "p-value", sig.level = -1,
         title = "VAS rating correlation matrix", 
         mar = c(0,0,1,0), 
         tl.cex=0.9) 
```

# Scatter plots
```{r scatter}
# Non TGI plots
ggplot(df_corr, aes(x = CNTcold, y =CNTwarm)) +
  geom_smooth(method='lm', colour = 'black', alpha = .5) +
  geom_point(size = 1.5, shape = 21, colour = 'grey10', fill = 'grey60') +
  geom_text(aes(label = paste0("R = ",
                               round(cormat[1,3], 2))),
            x=78, y=91, size = 5) +
  labs(x = 'Cold Ratings', y = 'Warm Ratings') +
  xlim(0,100) + ylim(-10,100) +
  theme_bw() +
  theme(axis.title = element_text(size = 12)) -> CNT_cw_plot

ggplot(df_corr, aes(x = CNTcold, y = CNTburn)) +
  geom_smooth(method='lm', colour = 'black', alpha = .5) +
  geom_point(size = 1.5, shape = 21, colour = 'grey10', fill = 'grey60') +
  geom_text(aes(label = paste0("R = ",
                               round(cormat[1,5], 2))),
            x=78, y=91, size = 5) +
  labs(x = 'Cold Ratings', y = 'Burn Ratings') +
  xlim(0,100) + ylim(-10,100) +
  theme_bw() +
  theme(axis.title = element_text(size = 12)) -> CNT_cb_plot

ggplot(df_corr, aes(x = CNTwarm, y = CNTburn)) +
  geom_smooth(method='lm', colour = 'black', alpha = .5) +
  geom_point(size = 1.5, shape = 21, colour = 'grey10', fill = 'grey60') +
  geom_text(aes(label = paste0("R = ",
                               round(cormat[3,5], 2))),
            x=78, y=91, size = 5) +
  labs(x = 'Warm Ratings', y = 'Burn Ratings') +
  xlim(0,100) + ylim(-10,100) +
  theme_bw() +
  theme(axis.title = element_text(size = 12)) -> CNT_wb_plot

# combine non TGI
CNT_corr_plot <- ggarrange(CNT_cw_plot, CNT_cb_plot, CNT_wb_plot,
                           nrow = 1, ncol = 3,
                           labels = c('A','B','C'))
CNT_corr_plot <- annotate_figure(CNT_corr_plot, 
                                 top = text_grob("Non-TGI", 
                                                 size = 14))

# TGI plots
ggplot(df_corr, aes(x = TGIcold, y = TGIwarm)) +
  geom_smooth(method='lm', colour = 'black', alpha = .5) +
  geom_point(size = 1.5, shape = 21, colour = 'grey10', fill = 'grey60') +
  geom_text(aes(label = paste0("R = ",
                               round(cormat[2,4], 2))),
            x=78, y=91, size = 5) +
  labs(x = 'Cold Ratings', y = 'Warm Ratings') +
  xlim(0,100) + ylim(-10,100) +
  theme_bw() +
  theme(axis.title = element_text(size = 12)) -> TGI_cw_plot

ggplot(df_corr, aes(x = TGIcold, y = TGIburn)) +
  geom_smooth(method='lm', colour = 'black', alpha = .5) +
  geom_point(size = 1.5, shape = 21, colour = 'grey10', fill = 'grey60') +
  geom_text(aes(label = paste0("R = ",
                               round(cormat[2,6], 2))),
            x=78, y=91, size = 5) +
  labs(x = 'Cold Ratings', y = 'Burn Ratings') +
  xlim(0,100) + ylim(-10,100) +
  theme_bw() +
  theme(axis.title = element_text(size = 12)) -> TGI_cb_plot

ggplot(df_corr, aes(x = TGIwarm, y = TGIburn)) +
  geom_smooth(method='lm', colour = 'black', alpha = .5) +
  geom_point(size = 1.5, shape = 21, colour = 'grey10', fill = 'grey60') +
  geom_text(aes(label = paste0("R = ",
                               round(cormat[4,6], 2))),
            x=78, y=91, size = 5) +
  labs(x = 'Warm Ratings', y = 'Burn Ratings') +
  xlim(0,100) + ylim(-10,100) +
  theme_bw() +
  theme(axis.title = element_text(size = 12)) -> TGI_wb_plot

# combine TGI
TGI_corr_plot <- ggarrange(TGI_cw_plot, TGI_cb_plot, TGI_wb_plot,
                           nrow = 1, ncol = 3,
                           labels = c('D','E','F'))
TGI_corr_plot <- annotate_figure(TGI_corr_plot, 
                                 top = text_grob("TGI", 
                                                 size = 14))

# put two together
corr_plot <- ggarrange(CNT_corr_plot, TGI_corr_plot,
                       nrow = 2, ncol = 1)
corr_plot

ggsave(here::here("figures", "E2_correlations.png"), plot = corr_plot, 
       device = NULL, width = 9, height = 5, dpi = 600)
```

# Initial burn
Are there any observed differences in dermatome and warm and cold afferents
for the first initial burning rating, before trial begins?

```{r initmod}
# Try and model all with ratings
model.init = glmmTMB::glmmTMB(beta ~ manipulation * condition * cold_cond + 
                                trial_n + (1|ID) + (1|order),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = df_long %>% filter(quality == 'init'),
                      na.action = na.omit) 
mi <- summary(model.init)
mi

```

# Run individual models on within and across dermatomes for H2
Directly compare proximal-distal and rostral-caudal results for each quality

# Cold - within and across
```{r cold_indiv}
# seperate data-frames
within_exp2 <- df_long_exp2 %>% 
  filter(condition == 'within')
across_exp2 <-  df_long_exp2 %>% 
  filter(condition == 'across')

# within
model.wcold = glmmTMB::glmmTMB(beta ~ manipulation * cold_probe + procedure + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = within_exp2 %>% filter(quality == 'cold'),
                      na.action = na.omit) 
mwc <- summary(model.wcold)
mwc

# across
model.acold = glmmTMB::glmmTMB(beta ~ manipulation * cold_probe + procedure + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = across_exp2 %>% filter(quality == 'cold'),
                      na.action = na.omit) 
mac <- summary(model.acold)
mac
```

# Warm - within and across
```{r warm_indiv}
# within
model.wwarm = glmmTMB::glmmTMB(beta ~ manipulation * cold_probe + procedure + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = within_exp2 %>% filter(quality == 'warm'),
                      na.action = na.omit) 
mww <- summary(model.wwarm)
mww

# across
model.awarm = glmmTMB::glmmTMB(beta ~ manipulation * cold_probe + procedure + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = across_exp2 %>% filter(quality == 'warm'),
                      na.action = na.omit) 
maw <- summary(model.awarm)
maw
```

# Burn - within and across
```{r burn_indiv}
# isolate responders only
within_exp2 <- df_long_exp2 %>% 
  filter(condition == 'within' & responder == 1)
across_exp2 <-  df_long_exp2 %>% 
  filter(condition == 'across' & responder == 1)

# within
model.wburn = glmmTMB::glmmTMB(beta ~ manipulation * cold_probe + procedure + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = within_exp2 %>% filter(quality == 'burn'),
                      na.action = na.omit) 
mwb <- summary(model.wburn)
mwb

# across
model.aburn = glmmTMB::glmmTMB(beta ~ manipulation * cold_probe + procedure + trial_n +
                                (1|ID),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+manipulation,
                      data = across_exp2 %>% filter(quality == 'burn'),
                      na.action = na.omit) 
mab <- summary(model.aburn)
mab
```